### 2016-06-20 paper/sn.tex
First commit

%\documentclass{article}
\documentclass[nofootinbib,amsmath,amssymb,10pt,eqsecnum, twocolumn]{revtex4-1}
\usepackage{graphicx, amsmath, amsmath, amssymb}
\usepackage[caption=false]{subfig}
\usepackage[breaklinks, colorlinks, citecolor=blue]{hyperref}
%\usepackage[maxbibnames=4]{biblatex}
\usepackage{color}
\usepackage{float}
\usepackage{natbib,hyperref,ifthen}
%------------ May need to comment out this -----------
%\usepackage{epstopdf}
%-----------less/greater than approx eq to--------------
\newcommand{\lta}{\mbox{\small\raisebox{-0.6ex}{$\,\stackrel
{\raisebox{-.2ex}{$\textstyle <$}}{\sim}\,$}}}
\newcommand{\gta}{\mbox{\small\raisebox{-0.6ex}{$\,\stackrel
{\raisebox{-.2ex}{$\textstyle >$}}{\sim}\,$}}}   
%-------------------------------------------------------
\newcommand{\edT}{\eta_{\Delta T}}
\newcommand{\nueff}{\nu_{\rm eff}}
\newcommand{\Tobs}{\tilde{T}_A^F}
\newcommand{\sigpix}{\sigma_{\rm pix}}
\newcommand{\sigsr}{\sigma_{\rm sr}}
\newcommand{\Tfa}{\Tobs}
\newcommand{\Tmod}{\hat{T}_A^F}
\newcommand{\Planck}{{\em Planck}}
\newcommand{\WMAP}{{\em WMAP}}
\newcommand{\un}{\, \rm}
\newcommand{\al}{\alpha}
\newcommand{\be}{\beta}
\newcommand{\ga}{\gamma}
\newcommand{\de}{\delta}
\newcommand{\s}{\sigma}
\newcommand{\pd}{\partial}
\newcommand{\brpd}{\bar{\partial}}
\newcommand{\Chr}[2]{\Gamma^{#1}{}_{\!#2}}
\newcommand{\brChr}[2]{\overline{\Gamma}^{#1}{}_{\!#2}}
\newcommand{\brg}{\bar{g}}
\newcommand{\brR}{\bar{R}}
\newcommand{\brG}{\bar{G}}
\newcommand{\brT}{\bar{T}}
\newcommand{\brLambda}{\overline\Lambda}
\newcommand{\brdel}{\overline\nabla}
\newcommand{\dd}{{\rm d}}
\newcommand{\jump}[1]{ \left[ #1 \right]^+_- }
\newcommand{\avrg}[1]{ \left\langle #1 \right\rangle^+_- } 
\newcommand{\Boxbar}{\overline{\Box}}
\newcommand{\diag}{\mbox{diag}}
\newcommand{\eps}{\varepsilon}
\newcommand{\oem}{{\cal T}}
\newcommand{\brW}{\overline{\cal W}}
\newcommand{\wave}{\square}
\newcommand{\brwave}{\overline{\square}}
\newcommand{\brkap}{{\bar\kappa}}
\newcommand{\bo}{\beta_0}
\newcommand{\binf}{\beta_{\!_\infty}}
\newcommand{\boldomega}{{\bf \omega}}
\newcommand{\appleq}{\raisebox{-3pt}{\makebox[14pt]{$\stackrel{\textstyle<}
{\sim}$}}}
\newcommand{\blockmatrix}[4]{\left( \begin{array}{c|c} #1
& #2 \\ \hline #3 & #4 \end{array} \right)}
\newcommand{\mbar}{{\bar m}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\btheta}{{\bf \hat{n}}}
\newcommand{\absmax}{{$|\mbox{Max}|$}}
\renewcommand{\arraystretch}{1.6}
\newcommand{\qsubrm}[2]{{#1}_{\scriptsize{\textrm{#2}}}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\bse{\begin{subequations}}

\bibliographystyle{arxiv}

\def\mnras{MNRAS}
\def\jcap{JCAP}

\newcommand\comment[1]{\textcolor{red}{#1}}

%-------------------------------------------------------

\begin{document}

\title{Deep Recurrent Neural Networks for Supernovae Classification}

\author{Tom Charnock} \email{tom.charnock@nottingham.ac.uk}
\affiliation{School of Physics \& Astronomy\\
University of Nottingham,
Nottingham, NG7 2RD, England}

\author{Adam Moss} \email{adam.moss@nottingham.ac.uk}
\affiliation{School of Physics \& Astronomy\\
University of Nottingham,
Nottingham, NG7 2RD, England}

\date{\today}

\begin{abstract}
   We apply deep recurrent neural networks, which are capable of learning complex sequential information, to classify supernovae\footnote{Code available at \href{https://github.com/adammoss/supernovae}{https://github.com/adammoss/supernovae}}. The observational time and  filter fluxes are used as inputs to the network, but since the inputs are agnostic additional data such as host galaxy information can also be included. Using the Supernovae Photometric Classification Challenge (SPCC) data, we find that deep networks are capable of learning about light curves, however the performance of the network is highly sensitive to the amount of training data.  For a training size of 50\% of the representational SPCC dataset (around $10^4$ supernovae) we obtain a type 1a vs non type 1a classification accuracy of 94.8\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$. We also apply a pre-trained model to obtain classification probabilities as a function of time, and find it can give early detections of supernovae type. Our method is competitive with existing algorithms and has applications for future large-scale photometric surveys. 
   \end{abstract}

\maketitle

\section{Introduction}

Future large, wide-field photometric surveys such as the Large Synoptic	Survey Telescope (LSST) will produce a vast amount of data, covering a large fraction of the sky every few nights. The amount of data produced lends itself to new analysis methods which can learn abstract representations of complex data. Deep learning is a powerful method for gaining multiple levels of abstraction, and has recently produced state-of-the-art results in tasks such as image classification and natural language processing. There are several types of deep learning architectures, such as convolutional neural networks, deep belief networks and recurrent neural networks.

There are many applications of deep learning for large photometric surveys, such as: (1) the measurement of galaxy shapes from images; (2) automated strong lens identification from multi-band images; (3) automated classification of supernovae; (4) galaxy cluster identification. In this paper we will focus on supernovae classification using deep recurrent neural networks. The LSST, for example, is expected to find over $10^7$ supernova~\cite{2009arXiv0912.0201L}. However,  it is estimated that only 5000 to 10,000\footnote{Although these numbers are not guaranteed.} will be spectroscopically confirmed  by follow up surveys~\cite{2013arXiv1311.2496M}, so classification methods need to be developed  for photometry.  All previous approaches to automated classification~\cite{Newling:2010bp, Karpenka:2012pm, Lochner:2016hbn} have first extracted features from supernovae light curves before using machine learning algorithms.  One of the advantages of deep learning is replacing this feature extraction.

{\em Recurrent neural networks} (RNNs) are a class of artificial neural network  that can learn about sequential data. They are commonly used for tasks such speech recognition and language translation, but have several possible applications in astronomy  and cosmology for processing temporal or spatial sequential data. RNNs have several properties which makes them suitable for sequential information. The inputs to the network are flexible, and they are able to recognise patterns with noisy data (for example the context of a word in a sentence relative to others can vary). 
\begin{figure}
\centering
\includegraphics[width=88mm, angle=0]{network.pdf}
\caption{\label{fig:network} Bidirectional recurrent neural network for sequence classification. The input vectors at each sequential step are fed into a pair of bidirectional hidden layers, which can propagate information forwards and backwards. These are then merged to obtain a consensus view of the network, and finally a softmax layer computes classification probabilities. 
 }
\end{figure}

The main problem with vanilla RNNs  is that they are unable to store long-term information, so inputs at the end of a sequence have no knowledge of inputs at the start. Several types of RNNs have been proposed to solve this problem, including {\em Long Short-Term Memory} (LSTM)~\cite{LSTM} and {\em Gated Recurrent Unit} (GRU)~\cite{2014arXiv1412.3555C} units. These are similar in concept, in that information is able to flow through the network via a gating mechanism. One further problem with RNNs is that information is only able to flow in one direction. In {\em bidirectional} RNNs information is able to pass both forwards and backwards. Bidirectional LSTM networks have been shown to be particularly powerful for supervised sequence labelling, where sequential data is accompanied by a set of discrete labels.

The architecture of a typical bidirectional RNN for sequence labelling is shown in Fig.~\ref{fig:network}, where the squares represent {\rm neurons}. In this case the inputs, which are vectors at each sequential step, are connected to two hidden RNN layers, which can be either vanilla RNN, LSTM or GRU units.  Each hidden layer contains a number of hidden units, and in each layer information flows either forwards or backwards,  but no information passes between the two directions. Several hidden layers can be stacked to form {\em deep} neural networks. Deep networks are capable of learning higher-level temporal or spatial representations, and complex relationships between the inputs and outputs.

The output from the final set of hidden layers in each direction is merged at each sequential step, and mean pooled (averaged) over all steps to obtain a consensus view of the network\footnote{We find that obtaining a consensus view improves the performance of the network.}. Finally, the mean output is fed to a {\em softmax} layer, which takes an input vector ${\bf z}$ and returns normalised, exponentiated outputs for each class label $i$, $\exp(z_i) / \sum_{i} \exp(z_i)$, i.e. a vector of probabilities.

 In the network, each neuron is connected to another by a weight matrix, and the optimal weights are found by back propagating the errors from a {\em loss function} of the output layer. For classification problems, this is typically the categorical cross-entropy between predictions and targets, defined as 
\begin{equation}
L= -\sum_{i,j} t_{i,j} \log \left( p_{i,j} \right)
\end{equation}
where $i,j$ run over the class labels, $t_{i,j}$ are the targets for each class (either 0 or 1) and $p_{i,j}$ are the predicted probabilities.

\section{Example Data}

In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, which consists of a set of 21319 simulated supernove light curves.  Each supernovae sample consists of a time series of flux measurements, with errors, in the $g, r, i, z$ bands (one band for each timestep), along with the position on the sky and dust extinction. An example set of light curves is shown in Fig.~\ref{fig:lightcurve}.

Due to the format of the input data, we first do a small amount of data processing to obtain values of the $g, r, i, z$ fluxes and errors at each sequential step.  First, for observations less than $\sim 1$ hour apart, we group theÊ $g, r, i, z$ values into a single vector. Observations are then of the form in Tab.~\ref{tab:augment}, where any missing values are denoted by a dash. In order to impute the missing value of $i$, we use {\em data augmentation} and randomly select a value between $i_1$ or $i_3$. We make 5 random augmentations of all missing data, thereby increasing the size of the dataset by a factor of 5.

\begin{table}[htdp]
\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline  
 Time & g & r & i & z \\ \hline \hline 
$ t_1$  & $g_1$ & $r_1$ &$ i_1$ & $z_1$ \\ \hline 
$ t_2$  & $g_2$ & $r_2$ &$  - $ & $z_2$ \\ \hline 
$ t_3$  & $g_3$ & $r_3$ &$ i_3$ & $z_3$ \\ \hline 
 \end{tabular}
\end{center}
\label{tab:augment}
\caption{Data augmentation of missing observations. The missing data is replaced randomly by a value between $i_1$ or $i_3$.}
\end{table}%

\begin{figure}
\centering
\includegraphics[width=88mm, angle=0]{DES_SN551675.pdf}
\caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g, r, i, z$ bands for SN ID 551675 (a type 1a) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}. The data has been processed using augmentation so there is a $g, r, i, z$ value at each sequential step. (Bottom) Type 1a probability as a function of time from a 2 layer LSTM model, trained with around $10^4$ supernovae and SN 551675 excluded. The final probability gives $99.5\%$ confidence that the supernovae is of type 1a. 
 }
\end{figure}

The data comes in two types, those with and those without the  host galaxy photometric redshift. Each dataset is split into a training and test set, with the training set containing a spectroscopically confirmed supernovae type and redshift. It is important that augmented data with the same supernovae ID go into either the training {\em or} test set, not a mix of both. The original SPCC data consisted of 1103 training samples. The answer keys were subsequently made available for the test set~\cite{Kessler:2010qj}.

The input vector to each sequential step then consists of: the time in days since the first observation; the flux in each of the 4 bands; the flux errors in each of the 4 bands; the RA and Dec; the dust extinction; and the host photo-z if relevant. Whilst we do not expect some of these variables to impact the classifier accuracy, we do not attempt any feature engineering and leave it to the network to decide if they are relevant.

RNNs typically perform better with more training data, so for training purposes we use the SPCC test set with answer keys (which is a non-biased representational dataset\footnote{The original SPCC training set was non-representational.}), and select a random fraction to act as the training set. We consider 1103 supernovae (a training fraction of 0.052), the same size as the original challenge, and a fraction of 0.5 (around $10^4$ supernovae), nearly an order of magnitude larger, and closer to the number likely to be spectroscopically followed up for LSST. The training performance of RNNs is also improved if the data is processed in mini-batches. In order to do this the input data must be of the same length, so we set the sequence length to be the maximum length over all supernovae observations, and prepend the input with padding. In training the network we ensure the padding is ignored by masking the padded input.

The goal of the classifier is to determine the supernovae type in the test set. We consider two problems, the first being to categorise two classes (type 1a vs non type 1a), and the second to categorise three classes (supernovae types 1, 2 and 3). We denote these as  `SN1a' and  `123' respectively.

We use several metrics to assess the classifier. The simplest is the accuracy, defined as the ratio between the number of correct predictions and total number of predictions. With two classes, a random classifier would have an accuracy of 0.5, and with three classes, an accuracy of 1/3.

Next are a variety of metrics coming from the {\em confusion matrix} of predictions.  For binary classification problems, the confusion matrix splits predictions into true positives (TP), false positives (FP),  false negatives (FN), and true negatives (TN). We consider the precision (also known as purity) and recall (also known as completeness) of the classifier. These are defined as 
\begin{equation}
{\rm Precision} = \frac{{\rm TP}}{{\rm TP}+{\rm FP}}\,, \quad {\rm Recall} = \frac{{\rm TP}}{{\rm TP}+{\rm FN}}\,.
\end{equation}
We evaluate these for each class separately vs `the rest' (e.g. type 1a vs non type 1a). The SPCC also defined the $F_1$ figure-of-merit for the SN1a classification problem. This is 
\begin{equation}
F_1 = \frac{1}{{\rm TP}+{\rm FN}} \frac{{\rm TP}^2}{{\rm TP}+3 \times {\rm FP}}\,,
\end{equation}
so false positives (incorrectly classifying a non type 1a supernovae as a type 1a) are penalised more heavily.

Finally, we also calculate  the Area Under the Curve (AUC). The AUC is the area under the curve of the TP rate vs FP rate, as the threshold probability for classification is increased from 0 to 1. A perfect classifier has an AUC of 1, and a random classifier 0.5. For multi-class problems, we calculate the AUC for each class vs the rest, and take an unweighted average to give the final AUC score.

\section{Network Architecture}

We consider several combinations of the network architecture. For the RNN type in the hidden layers, we test both vanilla RNN and long-term memory (LSTM and GRU) units. We also consider unidirectional  and bidirectional networks. For the unidirectional network, we fix the direction to be forwards. For the bidirectional network, we fix the number of hidden units in each RNN layer to be equal for the forward and backward directions.

We also test stacking two sets of layers to form a deep network.  In the unidirectional case we stack two hidden layers. In the bidirectional case the two stacks consists  of a pair  of forwards and backwards layers.  We denote the number of hidden units in a network with a  single stack by $[h_1]$, and the number of hidden layers in a two stack model by $[h_1, h_2]$. We vary the number of hidden units, testing h=[4], [8], [16], [32], [4,4], [8,8], [16,16] and [32,32]. We do not go beyond a stack of two layers due to the limited size of the dataset.

For each network we perform 5 randomised runs over the training data to obtain the classifier metrics. We define the loss function as the categorical cross-entropy between the predictions and test data. The network weights were trained using back-propogation with the `Adam' updater~\cite{2014arXiv1412.6980K}. Mini-batches of size 10\footnote{If training with a GPU larger mini-batches are recommended to make use of the GPU cores.} were used throughout, and each model was trained over approximately 200 {\em epochs}, where each epoch is a full pass over the training data.

\section{Results}

\begin{figure}
\centering
\includegraphics[width=88mm, angle=0]{lstm_dropout_05_loss.pdf}
\caption{\label{fig:loss} (Left) Training loss (green) vs test loss (blue) for a unidirectional 2 layer  LSTM network with 16 hidden units in each layer. (Right) Training accuracy (green) vs test accuracy (blue) for the same network.  }
\end{figure}

A dataset of 21319 is relatively small by deep learning standards. Furthermore, the `feature space' of supernovae light curves is significantly smaller than, say using RNNs to learn about language. We therefore need to be careful about {\em over-fitting}. Over-fitting arises when the network fits noise between the inputs and outputs of the training data, which do not exist in the test data. It can typically be detected by comparing the loss of the training and test data. If the loss of training data continues to decrease, but the loss of the test data increases, this is a sure sign of over-fitting. On the other hand, if no sign of over-fitting is observed, the network is not usually complex enough to fully learn the relationship between inputs and outputs (called {\em under-fitting}).

For a training fraction of 0.5, we found the best architecture was a deep 2 layer network with unidirectional LSTM units. Bidirectional units did not significantly improve the test accuracy and made the network more difficult to train. There was a marked improvement in test accuracy using 16 hidden units in each layer rather than 8, but too much over-fitting occured using 32 hidden units. Over-fitting was still an issue for 16 hidden units,  but we make use of a technique called {\em dropout}~\cite{JMLR:v15:srivastava14a} to regularize this. Dropout  sets a random fraction of connections to 0 at each update during training {\em only}, which prevents the units from adapting too much. We apply dropout only to non-recurrent connections after each hidden layer.

In Fig.~\ref{fig:loss} we show the training and tests losses for such a network, with a dropout of $0.5$, applied to type 1a vs non type 1a classification with host galaxy photo-z information. Without dropout the training loss would continue to fall and the test loss would rise. For 5 randomised runs, after training for  200 epochs, we obtain a classification accuracy of $94.8 \pm 0.2$\%, an AUC of $0.986 \pm 0.001$ and $F_1 = 0.64 \pm 0.01$. The corresponding type 1a purity and completeness are 87\% and 92\% respectively. A summary of results for this network can be found in Tabs.~\ref{tab:05_SN1a} and~\ref{tab:05_123}. It can be seen that the inclusion of host galaxy photo-z does not significantly improve the classifier performance.

\begin{table}[t!]
\begin{tabular}{|c | c | c | c | c |} \hline 
AUC & Accuracy & F1 & Host z \\ \hline 
$ 0.986 \pm  0.001$ &  $  94.8 \pm 0.2\% $ & $  0.64 \pm 0.01  $ & T  \\
$ 0.981 \pm  0.002$ &  $  93.6 \pm 0.3\% $ & $  0.60 \pm 0.02  $ & F  \\ \hline
\end{tabular}
\caption{\label{tab:05_SN1a} Summary of results for  type 1a vs non type 1a classification with a training fraction of 0.5. The network used is a 2 layer LSTM, each with 16 hidden units, and a dropout of 0.5.}
\end{table}

\begin{table}[t!]
\begin{tabular}{|c | c | c | c | c |} \hline 
AUC & Accuracy & Host z  \\ \hline 
$ 0.971  \pm 0.003 $ &  $  90.4  \pm 0.4 \% $ & T  \\
$ 0.958 \pm  0.007 $ &  $   88.8 \pm 1.0 \% $ & F \\ \hline
\end{tabular}
\caption{\label{tab:05_123} Summary of results for three class categorisation (supernovae types 1, 2 and 3). The model is the same as in Tab.~\ref{tab:05_SN1a}}
\end{table}
For $5.2\%$ of the representative SPCC data, the training dataset is so small that over-fitting is more severe. Using the same 2 layer LSTM network with 16 hidden units and a dropout of 0.5 we find a notable increase in the test loss after $\sim 20$ epochs, but the accuracy and other metrics remain relatively constant ($F_1$ values, for example, of 0.35 to 0.4 were obtained). The reason for this apparent discrepancy is that the accuracy, say, simply takes the maximum value of the softmax output layer. For example, a 2 class problem with output probabilities [0.6, 0.4] and target [1, 0] has the same accuracy as one with output probabilities [0.8, 0.2]. The loss in the latter case would be lower however, and represents increased confidence of the network in its predictions. We therefore reject models with over-fitting and an increasing cross-entropy loss, at the expense of metrics such as $F_1$, and decrease the model complexity.

For a training fraction of  $5.2\%$ we find a single layer LSTM network, with 4 hidden units, and a dropout of 0.5 satisfies this criteria. For 5 randomised runs, after training for  200 epochs, we obtain a classification accuracy of $85.9 \pm 1.2$\%, an AUC of $0.910 \pm 0.012$ and $F_1 = 0.31 \pm 0.03$. The corresponding type 1a purity and completeness are 72\% and 66\% respectively.

\section{Conclusions}

We have presented a new method for performing  photometric classification of supernovae. Machine learning methodology has previously been applied to SPCC classification~\cite{Newling:2010bp, Karpenka:2012pm, Lochner:2016hbn}. These all employ a two step process, where features are first extracted by some method before classification. In our approach light curves are used directly as inputs to a recurrent neural network, which is able to learn information from the sequence of observations.

Using post SPCC data, we obtain results which are competitive with previous approaches.  Although we have trained the network on the cross-entropy loss  and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-z), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.65). Recurrent neural networks  therefore compare very well with a larger training  set. The performance isn't  quite as good with a smaller training set, possibly due to the network having to learn from no prior information about (noisy) light curves. This is also seen in~\cite{Lochner:2016hbn}, which obtains improved metrics  for only a small training fraction with a combination of feature extraction and  machine learning.

One advantage of our approach is that light curve data can be directly inputted to a pre-trained model to give very fast evaluation ($<1\,$s) of supernovae type. In the lower panel of Fig.~\ref{fig:lightcurve} we input the light curve of a type 1a supernovae (excluded from training) to a pre-trained 2 layer LSTM model with a training fraction of 0.5. The classifier (type 1a vs non type 1a) is initially unsure of classification, with a type 1a probability of around 0.5. The probability then decreases slightly, but rapidly increases near the peak of the light curve. The classifier has high confidence the supernovae is of type 1a at around 60 days, and the final probability is excess of $99.5\%$.  This method could therefore be useful for real time monitoring in  surveys.

There are several possibilities for future work. One of the advantages of recurrent neural networks is that inputs are agnostic, so the impact of any additional inputs could  be explored. It would be possible, for example, to even pass the raw images in each filter though a convolutional network and use those as inputs. We have considered a representative training sample, but spectroscopic follow up surveys may be biased. The performance of the network could be measured against selection bias, and the results used to inform the best follow up strategy.  Further work could also be performed to optimise the early detection probability of the network.  Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is be pre-trained on simulations or existing data from other surveys, then the weights of the network are fine-tuned on the new, smaller dataset.

\section*{Acknowledgements}

We appreciate helpful conversations with Steven Bamford, Simon Dye, Mark Sullivan and Michael Wood-Vasey, and Natasha Karpenka, Richard Kessler and Michelle Lochner for help with data acquisition. T. C. is supported by a STFC studentship, and A. M. is supported by a Royal Society University Research Fellowship.

\bibliography{bib}

\end{document}

### 2016-06-20 arch.py

@@ -1,0 +1,182 @@
+from train import *
+import itertools as it
+from multiprocessing import Pool
+import json
+import numpy as np
+from random import choice
+from string import ascii_uppercase
+import glob
+import argparse
+
+nb_rand = 5
+root_filename = 'info'
+
+def analyse(filename):
+
+	results = []
+	maps = []
+
+	for f in glob.glob('arch/'+filename+'*.txt'):
+		with open(f) as data_file:   
+			p = json.load(data_file)
+			if 'auc' in p:
+				try:
+					results.append([np.mean(p['auc']), p['classifier'][1], p, f])
+					maps.append(p['classifier'][1])
+				except:
+					results.append([np.mean(p['auc']), p['mapping'][1], p, f])
+					maps.append(p['mapping'][1])
+
+	for m in list(set(maps)):
+
+		map_results = [r for r in results if r[1] == m]
+		sorted_results = sorted(map_results, key=lambda x: -x[0])
+
+		print('classifier: ', m)
+
+		with open("arch/"+filename+"_"+m+".tex", "w") as f:
+			f.write("\\begin{tabular}{|" + " | ".join(["c"] * 9) + "|} \hline \n")
+			f.write(" & ".join(['AUC', 'Accuracy', 'F1', 'z', 'Class', 'h', 'B']) + " \\\\ \hline \n")
+			for row in sorted_results:
+				p = row[2]
+				acc_mean = np.mean(p['acc'])
+				acc_std = np.std(p['acc'])
+				auc_mean = np.mean(p['auc'])
+				auc_std = np.std(p['auc'])
+				f1_mean = np.mean(p['f1'])
+				f1_std = np.std(p['f1'])
+				prec = np.array(p['prec'])
+				rec = np.array(p['rec'])
+				prec_mean = [ "$ {:.3f}".format(np.mean(prec[:,i])) for i in range(0,prec.shape[1])]
+				prec_std = [ "$ {:.3f}".format(np.std(prec[:,i])) for i in range(0,prec.shape[1])]
+				rec_mean = [ "$ {:.3f}".format(np.mean(rec[:,i])) for i in range(0,prec.shape[1])]
+				rec_std = [ "$ {:.3f}".format(np.std(rec[:,i])) for i in range(0,prec.shape[1])]
+				auc = "$ {:.3f}".format(auc_mean) + ' \pm ' + "{:.3f} $".format(auc_std)
[…]

### 2016-06-20 custom_layers.py

@@ -1,0 +1,525 @@
+from keras import backend as K
+from keras.engine import Layer
+from keras.layers.core import Lambda
+from keras.layers import Merge
+import theano
+
+import copy
+import inspect
+import types as python_types
+import marshal
+import sys
+import warnings
+
+class MaskLambda(Layer):
+  
+    def __init__(self, function, output_shape=None, arguments={}, **kwargs):
+        self.function = function
+        self.arguments = arguments
+        self.supports_masking = True
+        if output_shape is None:
+            self._output_shape = None
+        elif type(output_shape) in {tuple, list}:
+            self._output_shape = tuple(output_shape)
+        else:
+            if not hasattr(output_shape, '__call__'):
+                raise Exception('In Lambda, `output_shape` '
+                                'must be a list, a tuple, or a function.')
+            self._output_shape = output_shape
+        super(MaskLambda, self).__init__(**kwargs)
+
+    def get_output_shape_for(self, input_shape):
+        if self._output_shape is None:
+            # if TensorFlow, we can infer the output shape directly:
+            if K._BACKEND == 'tensorflow':
+                if type(input_shape) is list:
+                    xs = [K.placeholder(shape=shape) for shape in input_shape]
+                    x = self.call(xs)
+                else:
+                    x = K.placeholder(shape=input_shape)
+                    x = self.call(x)
+                if type(x) is list:
+                    return [K.int_shape(x_elem) for x_elem in x]
+                else:
+                    return K.int_shape(x)
+            # otherwise, we default to the input shape
+            return input_shape
+        elif type(self._output_shape) in {tuple, list}:
+            nb_samples = input_shape[0] if input_shape else None
+            return (nb_samples,) + tuple(self._output_shape)
+        else:
+            shape = self._output_shape(input_shape)
+            if type(shape) not in {list, tuple}:
+                raise Exception('output_shape function must return a tuple')
+            return tuple(shape)
+
+    def call(self, x, mask=None):
+        arguments = self.arguments
+        arg_spec = inspect.getargspec(self.function)
+        if 'mask' in arg_spec.args:
+            arguments['mask'] = mask
+        return self.function(x, **arguments)
+
+    def get_config(self):
+        py3 = sys.version_info[0] == 3
+
+        if isinstance(self.function, python_types.LambdaType):
+            if py3:
+                function = marshal.dumps(self.function.__code__).decode('raw_unicode_escape')
[…]

### 2016-06-20 dataset.py

@@ -1,0 +1,208 @@
+from __future__ import print_function
+import numpy as np
+import csv
+
+sn1a_classifier = {'1':0, '2':1, '3':1, '21':1, '22': 1, '23':1, '32': 1, '33':1}
+type_classifier = {'1':0, '2':1, '3':2, '21':1, '22': 1, '23':1, '32': 2, '33':2}
+subtype_classifier = {'1':0, '21':1, '22':2, '23':3, '32': 4, '33':5, '3':6}
+
+def pad_sequences(sequences, maxlen=None, dtype='int32',
+				  padding='pre', truncating='pre', value=-1.):
+	'''
+	Pads each sequence to the same length, which is the length of the longest sequence. Returns
+        a numpy array of the padded sequences. 
+	* maxlen is an integer which truncates any sequence longer than maxlen and fills
+          any sequence less than maxlen up to the value of maxlen
+	* dtype is a string containing the type to cast the resulting sequence
+	* padding is a string, either "pre" or "post" which dictates whether to pad at the start
+	  or end of the sequence
+	* truncating is a string, either "pre" or "post" which dictates whether to truncate at 
+          the start or end of the sequence
+	* value is a float containing the value to pad the sequences with
+	* x is the numpy array containing the padded sequences with with dimensions 
+	  (number_of_sequences, maxlen)
+	- Used in load_data() when padding the training and test data
+	- Used in predict_probability() in train.py to pad the light curve data
+	'''
+	lengths = [len(s) for s in sequences]
+
+	nb_samples = len(sequences)
+	if maxlen is None:
+		maxlen = np.max(lengths)
+
+	# take the sample shape from the first non empty sequence
+	# checking for consistency in the main loop below.
+	sample_shape = tuple()
+	for s in sequences:
+		if len(s) > 0:
+			sample_shape = np.asarray(s).shape[1:]
+			break
+
+	x = (np.ones((nb_samples, maxlen) + sample_shape) * value).astype(dtype)
+	for idx, s in enumerate(sequences):
+		if len(s) == 0:
+			continue  # empty list was found
+		if truncating == 'pre':
+			trunc = s[-maxlen:]
+		elif truncating == 'post':
+			trunc = s[:maxlen]
+		else:
+			raise ValueError('Truncating type "%s" not understood' % truncating)
+
+		# check `trunc` has expected shape
+		trunc = np.asarray(trunc, dtype=dtype)
+		if trunc.shape[1:] != sample_shape:
[…]

### 2016-06-20 preprocess.py

@@ -1,0 +1,470 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import glob
+import csv
+import gzip
+import scipy.interpolate as si
+from itertools import groupby
+import random
+import sys
+import argparse
+
+flux_norm = 1.0
+time_norm = 1.0
+position_norm = 1.0
+grouping = 1
+
+key_types = {'Ia':1, 'II': 2, 'Ibc': 3, 'IIn': 21, 'IIP': 22, 'IIL': 23, 'Ib': 32, 'Ic': 33}
+
+def index_min(values):
+	'''
+	Return the index of an array.
+	* values is an array (intended to be times)
+	- Used in time_collector() for grouping times
+	- Used in parser_spline() for placing flux errors at the correct time in the time sequence
+	'''
+	return min(xrange(len(values)),key=values.__getitem__)
+
+def spline(arr,t):
+	'''
+	Returns the interpolated spline value of the fluxes at a given time. If the length of the
+	array is too short for the spline routine it pads the values so that spline interpolation
+	can be carried out.
+	* arr is an array with arr = [[times],[fluxes],[flux errors]]
+	* t is array of grouped times
+	- Used in parser_spline() to create the spline of the fluxes
+	''' 
+	if (len(arr[0]) < 4):
+		if (len(arr[0])==0):
+			arr[0] = [t[0],int((t[-1]-t[0])/2),t[-1]]
+			arr[1] = [0,0,0] 
+			arr[2] = [0,0,0] 
+		if (len(arr[0])==1):
+			arr[0] = [t[0],arr[0][0],t[-1]]
+			arr[1] = [arr[1][0],arr[1][0],arr[1][0]] 
+			arr[2] = [arr[2][0],arr[2][0],arr[2][0]] 
+		spline = si.InterpolatedUnivariateSpline(arr[0], arr[1],k=1)
+	else:
+		spline = si.InterpolatedUnivariateSpline(arr[0], arr[1])	
+	return spline
+
+def time_collector(arr,frac=grouping):
+	'''
+	Returns the an array of average times about clustered observation times. Default grouping is
+	for times on the order of 1 day, although this is reduced if there are too many observations
+	in that time. Also returns the index of the indices of the closest times in each flux band
+	and the grouping fraction.
+	* arr is an array containing all of the observation times
+	* frac is the clustering scale where frac=1 is group times within a day
[…]

### 2016-06-20 run.py

@@ -1,0 +1,63 @@
+from train import *
+import time
+import ConfigParser
+import argparse
+import json
+import itertools as it
+import os.path
+
+if __name__ == '__main__':
+
+	parser = argparse.ArgumentParser(description='')
+	parser.add_argument('-f','--f', type=str, help='Ini filename')
+	args = parser.parse_args()
+
+	if args.f:
+		filename = args.f
+	else:
+		filename = 'test.ini'
+
+	if not os.path.isfile(filename):
+		raise Exception('File does not exist') 
+
+	config = ConfigParser.ConfigParser()
+	config.read(filename)
+
+	rnn_type = config.get('Network','rnn_type')
+	dropout = config.getfloat('Network','dropout')
+	nb_hidden = json.loads(config.get('Network','hidden_layers'))
+	bidirectional = config.getboolean('Network','bidirectional')
+	consensus = config.getboolean('Network','consensus')
+
+	optimizer = config.get('Train','optimizer')
+	test_fraction = config.getfloat('Train','test_fraction')
+	nb_epoch = config.getint('Train','num_epochs')
+	batch_size = config.getint('Train','batch_size')
+
+	nb_augment = config.getint('Data','num_augment')
+	path = config.get('Data','data')
+	classifier_type = config.get('Data','classifier')
+
+	if classifier_type == 'SN1a':
+		classifier = sn1a_classifier
+	elif classifier_type == '123':
+		classifier = type_classifier
+	elif classifier_type.lower() == 'Sub':
+		classifier = subtype_classifier
+	else:
+		raise Exception('Incorrect classifier') 
+
+	root_filename = config.get('Options','filename')
+	plot_loss = config.getboolean('Options','plot_loss')
+	save_model = config.getboolean('Options','save_model')
+
+	with open('save/' + root_filename + '.ini', 'w') as f:
+		config.write(f)
+
+	start_time = time.time()
+	result = train(batch_size=batch_size, dropout=dropout, nb_hidden=nb_hidden, path=path, test_fraction=test_fraction, 
+			classifier=classifier, nb_epoch=nb_epoch, bidirectional=bidirectional, rnn_type=rnn_type, save_model=save_model, 
+			plot_loss=plot_loss, filename=root_filename, optimizer=optimizer, nb_augment=nb_augment,
+			consensus=consensus)
+	print("--- %s seconds ---" % (time.time() - start_time))
+

### 2016-06-20 train.py

@@ -1,0 +1,366 @@
+from __future__ import print_function
+from dataset import *
+from preprocess import parser_augment, parser_last, parser_spline, key_types
+from utils import loss_plotter, lightcurve_plotter
+import numpy as np
+import pickle
+
+def build_model(nb_hidden, nb_classes, dropout, sequence_len, output_dim, bidirectional=False, rnn_type='LSTM', activation='tanh', consensus=True): 
+
+	'''
+	Builds model of the neural network. Bidirectional layers are implemented as 2 forward 
+        layers with data reversed in 1 layer to align the masks
+	* nb_hidden is list of integers with the number of hidden units, e.g. [16, 16]
+	* nb_classes is an integer with the number of classification classes
+	* dropout is a float between 0 and 1 describing dropout probability after each hidden 
+          layer (non-recurrent connections only)
+	* sequence_len is an integer containing the maximum length of observations
+	* output_dim is an integer containing length of vector for each step
+	* bidirectional is a boolean which when True uses a bidirectional network
+	* rnn_type is a string giving RNN type (either "RNN", "GRU", "LSTM")
+	* activation is a string describing the activation function for the hidden layer
+	* consensus is a boolean which when true mean pools the predictions from each step
+	- Used in train() to build the neural network for training
+	- Used in predict_probability() to build the neural network for analysis using the
+	  weights saved from the training
+	'''
+
+	import keras.backend as K
+	from keras.preprocessing import sequence
+	from keras.models import Sequential, Model
+	from keras.layers.core import Lambda, Masking
+	from keras.layers import Dense, Dropout, Activation, Embedding, Merge, Input
+	from keras.layers import LSTM, SimpleRNN, GRU
+	from keras.layers.wrappers import TimeDistributed
+	from custom_layers import MaskLambda, lambda_mask_average, MaskMerge, mask_merge
+
+	if rnn_type.lower() == 'lstm':
+		RNNLayer = LSTM
+	elif rnn_type.lower() == 'gru':
+		RNNLayer = GRU
+	else:
+		RNNLayer = SimpleRNN
+
+	if bidirectional:
+
+		input_a = Input(shape=(sequence_len, output_dim))
[…]

### 2016-06-20 utils.py

@@ -1,0 +1,82 @@
+'''
+Plotting routines
+'''
+
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import numpy as np
+
+def loss_plotter(history, filename, plot_legend=True):
+	'''
+	Plots the test and training loss history and accuracy history after every
+	epoch of training. Plots are saved in ./arch/
+	* history is a list which contains the training loss and test loss and the training
+	  accuracy and test accuracy which are each contained in the array. Each random
+	  augmentation of train() is another dimension in the array.
+	* filename is the string containing the filename to name the plot
+	* plot_legend is a boolean which when True places a legend on each axis
+	- Used in train(), particularly it is in on_epoch_end() from LossHistory() which is a callback
+	  called at the end of each epoch of training
+	'''
+	plt.rc('legend',**{'fontsize':16})
+	fig = plt.figure()
+	gs = gridspec.GridSpec(1,2)
+	gs.update(left=0.1,right=0.97,wspace=0.3)
+	ax1 = plt.subplot(gs[0])
+	ax1.set_xlabel('Epoch',fontsize=16)
+	ax1.set_ylabel('Loss',fontsize=16,labelpad=0)
+	ax1.xaxis.grid(True,'minor')
+	ax1.yaxis.grid(True,'minor')
+	ax1.xaxis.grid(True,'major')
+	ax1.yaxis.grid(True,'major')
+	plt.tick_params(axis='both',which='major',labelsize=16)
+	ax2 = plt.subplot(gs[1])
+	ax2.set_xlabel('Epoch',fontsize=16)
+	ax2.set_ylabel('1 - Accuracy',fontsize=16,labelpad=0)
+	ax2.xaxis.grid(True,'minor')
+	ax2.yaxis.grid(True,'minor')
+	ax2.xaxis.grid(True,'major')
+	ax2.yaxis.grid(True,'major')
+	plt.tick_params(axis='both',which='major',labelsize=16)
+	for i in xrange(len(history)):	
+		if i < len(history)-1:
+			ax1.semilogy(history[i][0],color='green')
+			ax1.semilogy(history[i][1],color='blue')
+			ax2.semilogy(1 - np.array(history[i][2]),color='green')
+			ax2.semilogy(1 - np.array(history[i][3]),color='blue')
+		else:
+			a, = ax1.semilogy(history[i][0],color='green')
+			b, = ax1.semilogy(history[i][1],color='blue')
[…]

### 2016-06-20 paper/sn.tex
Typo

@@
 Further work could also be performed to optimise the early detection probability of the network.
-Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is be pre-trained on simulations or existing data from other surveys, then the weights of the network are fine-tuned on the new, smaller dataset.
+Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is pre-trained on simulations or existing data from other surveys, then the weights of the network are fine-tuned on the new, smaller dataset.
 

### 2016-06-21 paper/sn.tex
Paper update

@@
 Using the Supernovae Photometric Classification Challenge (SPCC) data, we find that deep networks are capable of learning about light curves, however the performance of the network is highly sensitive to the amount of training data.
-For a training size of 50\% of the representational SPCC dataset (around $10^4$ supernovae) we obtain a type 1a vs non type 1a classification accuracy of 94.8\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$.
-We also apply a pre-trained model to obtain classification probabilities as a function of time, and find it can give early detections of supernovae type.
+For a training size of 50\% of the representational SPCC dataset (around $10^4$ supernovae) we obtain a type Ia vs non type Ia classification accuracy of 94.8\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$.
+We also apply a pre-trained model to obtain classification probabilities as a function of time, and show it can give early indications of supernovae type.
 Our method is competitive with existing algorithms and has applications for future large-scale photometric surveys.
@@
 Deep learning is a powerful method for gaining multiple levels of abstraction, and has recently produced state-of-the-art results in tasks such as image classification and natural language processing.
-There are several types of deep learning architectures, such as convolutional neural networks, deep belief networks and recurrent neural networks.
+There are several types of deep learning architectures, such as convolutional neural networks, deep belief networks and recurrent neural networks (see~\cite{0483bd9444a348c8b59d54a190839ec9} for an excellent overview of deep learning and refs.
+within for more details).
 
@@
 
-{\em Recurrent neural networks} (RNNs) are a class of artificial neural network  that can learn about sequential data.
-They are commonly used for tasks such speech recognition and language translation, but have several possible applications in astronomy  and cosmology for processing temporal or spatial sequential data.
-RNNs have several properties which makes them suitable for sequential information.
-The inputs to the network are flexible, and they are able to recognise patterns with noisy data (for example the context of a word in a sentence relative to others can vary).
+In this work we will use {\em supervised} deep learning.
+During training, the machine is given inputs and produces a set of output predictions.
+It is also given the correct set of outputs.
+An objective loss function then measures the error between the predicted and target outputs, and the machine updates its adjustable parameters to reduce the error.
+It can then make predictions for unknown outputs.
+
 \begin{figure}
@@
 
+{\em Recurrent neural networks} (RNNs) are a class of artificial neural network  that can learn about sequential data.
+They are commonly used for tasks such speech recognition and language translation, but have several possible applications in astronomy  and cosmology for processing temporal or spatial sequential data.
+RNNs have several properties which makes them suitable for sequential information.
+The inputs to the network are flexible, and they are able to recognise patterns with noisy data (for example the context of a word in a sentence relative to others can vary).
+
 The main problem with vanilla RNNs  is that they are unable to store long-term information, so inputs at the end of a sequence have no knowledge of inputs at the start.
@@
 The architecture of a typical bidirectional RNN for sequence labelling is shown in Fig.~\ref{fig:network}, where the squares represent {\rm neurons}.
-In this case the inputs, which are vectors at each sequential step, are connected to two hidden RNN layers, which can be either vanilla RNN, LSTM or GRU units.
+In this case the inputs, which are vectors at each sequential step, are connected to two hidden RNN layers, which can be either vanilla RNN or memory units.
 Each hidden layer contains a number of hidden units, and in each layer information flows either forwards or backwards,  but no information passes between the two directions.
@@
 Due to the format of the input data, we first do a small amount of data processing to obtain values of the $g, r, i, z$ fluxes and errors at each sequential step.
-First, for observations less than $\sim 1$ hour apart, we group theÊ $g, r, i, z$ values into a single vector.
+First, we assume the time sequence begins at day 0 for each supernovae, rather than counting days forwards and backwards from the maxima of the light curve.
+Next, for observations less than $\sim1$ hour apart, we group the $g, r, i, z$ values into a single vector, ensuring there is at most one filter type in each group.
+If there is more than one filter type, we further subdivide the group using a finer time interval.
+We take the time of the group to be the mean time of its individual observations, which is reasonable as the time intervals are small compared to the characteristic time of the light curve.
+
  Observations are then of the form in Tab.~\ref{tab:augment}, where any missing values are denoted by a dash.
@@
 We make 5 random augmentations of all missing data, thereby increasing the size of the dataset by a factor of 5.
+We can test the importance of augmentation by training each augmentation separately and  comparing the change in accuracy.
+We find that augmentation gives a percent level improvement in accuracy.
 
@@
 \includegraphics[width=88mm, angle=0]{DES_SN551675.pdf}
-\caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g, r, i, z$ bands for SN ID 551675 (a type 1a) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}.
+\caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g, r, i, z$ bands for SN ID 551675 (a type Ia) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}.
 The data has been processed using augmentation so there is a $g, r, i, z$ value at each sequential step.
-(Bottom) Type 1a probability as a function of time from a 2 layer LSTM model, trained with around $10^4$ supernovae and SN 551675 excluded.
-The final probability gives $99.5\%$ confidence that the supernovae is of type 1a.
+(Bottom) Type Ia probability as a function of time from a 2 layer LSTM model, trained with around $10^4$ supernovae and SN 551675 excluded.
+The final probability gives $99.5\%$ confidence that the supernovae is of type Ia.
  }
@@
 RNNs typically perform better with more training data, so for training purposes we use the SPCC test set with answer keys (which is a non-biased representational dataset\footnote{The original SPCC training set was non-representational.}), and select a random fraction to act as the training set.
-We consider 1103 supernovae (a training fraction of 0.052), the same size as the original challenge, and a fraction of 0.5 (around $10^4$ supernovae), nearly an order of magnitude larger, and closer to the number likely to be spectroscopically followed up for LSST.
+We consider 1103 supernovae (a training fraction of 0.052), the same size as the original challenge, and fractions of 0.25 and 0.5 (around 5000 and $10^4$ supernovae respectively), nearly an order of magnitude larger, and closer to the number likely to be followed up for the LSST.
 The training performance of RNNs is also improved if the data is processed in mini-batches.
@@
 The goal of the classifier is to determine the supernovae type in the test set.
-We consider two problems, the first being to categorise two classes (type 1a vs non type 1a), and the second to categorise three classes (supernovae types 1, 2 and 3).
+We consider two problems, the first being to categorise two classes (type Ia vs non type Ia), and the second to categorise three classes (supernovae types 1, 2 and 3).
 We denote these as  `SN1a' and  `123' respectively.
@@
 For binary classification problems, the confusion matrix splits predictions into true positives (TP), false positives (FP),  false negatives (FN), and true negatives (TN).
-We consider the precision (also known as purity) and recall (also known as completeness) of the classifier.
+We consider the purity (also known as precision) and completeness (also known as recall) of the classifier.
 These are defined as
 \begin{equation}
-{\rm Precision} = \frac{{\rm TP}}{{\rm TP}+{\rm FP}}\,, \quad {\rm Recall} = \frac{{\rm TP}}{{\rm TP}+{\rm FN}}\,.
+{\rm Purity} = \frac{{\rm TP}}{{\rm TP}+{\rm FP}}\,, \quad {\rm Completeness} = \frac{{\rm TP}}{{\rm TP}+{\rm FN}}\,.
 \end{equation}
-We evaluate these for each class separately vs `the rest' (e.g. type 1a vs non type 1a). The SPCC also defined the $F_1$ figure-of-merit for the SN1a classification problem. This is 
+We evaluate these for each class separately vs `the rest' (e.g. type Ia vs non type Ia). The SPCC also defined the $F_1$ figure-of-merit for the SN1a classification problem. This is 
 \begin{equation}
@@
 \end{equation}
-so false positives (incorrectly classifying a non type 1a supernovae as a type 1a) are penalised more heavily.
+so false positives (incorrectly classifying a non type Ia supernovae as a type Ia) are penalised more heavily.
 
@@
 
-In Fig.~\ref{fig:loss} we show the training and tests losses for such a network, with a dropout of $0.5$, applied to type 1a vs non type 1a classification with host galaxy photo-z information.
+In Fig.~\ref{fig:loss} we show the training and tests losses for such a network, with a dropout of $0.5$, applied to type Ia vs non type Ia classification with host galaxy photo-z information.
 Without dropout the training loss would continue to fall and the test loss would rise.
 For 5 randomised runs, after training for  200 epochs, we obtain a classification accuracy of $94.8 \pm 0.2$\%, an AUC of $0.986 \pm 0.001$ and $F_1 = 0.64 \pm 0.01$.
-The corresponding type 1a purity and completeness are 87\% and 92\% respectively.
+The corresponding type Ia purity and completeness are 87\% and 92\% respectively.
 A summary of results for this network can be found in Tabs.~\ref{tab:05_SN1a} and~\ref{tab:05_123}.
@@
 \end{tabular}
-\caption{\label{tab:05_SN1a} Summary of results for  type 1a vs non type 1a classification with a training fraction of 0.5.
+\caption{\label{tab:05_SN1a} Summary of results for  type Ia vs non type Ia classification with a training fraction of 0.5.
 The network used is a 2 layer LSTM, each with 16 hidden units, and a dropout of 0.5.}
@@
 \end{table}
+
+One advantage of our approach is that light curve data can be directly inputted to a {\em pre-trained} model to give very fast evaluation ($<1\,$s) of supernovae type.
+In the lower panel of Fig.~\ref{fig:lightcurve} we input the light curve, as a function of time, of a type Ia supernovae (excluded from training) to the pre-trained 2 layer LSTM model discussed above.
+The classifier (type Ia vs non type Ia) is initially unsure of classification, with a type Ia probability of around 0.5.
+The probability then decreases slightly, but rapidly increases near the peak of the light curve.
+The classifier has high confidence the supernovae is of type Ia at around 60 days, and the final probability is excess of $99.5\%$.
+This method could therefore be useful to give early indication of supernovae type in  surveys.
+
+We also test the same model using a training fraction of 0.25, around 5000 supernovae, close to the lower  end of the number likely to be followed up for the LSST.
+Again, after 5 randomised runs and training for 200 epochs the accuracy was \comment{$***\pm***$}, the AUC was \comment{$***\pm***$} and \comment{$F_1 =***\pm***$}.
+This suggests that the classifer metrics scale with the data \comment{$\propto***$}.
+
 For $5.2\%$ of the representative SPCC data, the training dataset is so small that over-fitting is more severe.
@@
 For 5 randomised runs, after training for  200 epochs, we obtain a classification accuracy of $85.9 \pm 1.2$\%, an AUC of $0.910 \pm 0.012$ and $F_1 = 0.31 \pm 0.03$.
-The corresponding type 1a purity and completeness are 72\% and 66\% respectively.
+The corresponding type Ia purity and completeness are 72\% and 66\% respectively.
 
@@
 This is also seen in~\cite{Lochner:2016hbn}, which obtains improved metrics for only a small training fraction with a combination of feature extraction and machine learning.
-
-One advantage of our approach is that light curve data can be directly inputted to a pre-trained model to give very fast evaluation ($<1\,$s) of supernovae type.
-In the lower panel of Fig.~\ref{fig:lightcurve} we input the light curve of a type 1a supernovae (excluded from training) to a pre-trained 2 layer LSTM model with a training fraction of 0.5.
-The classifier (type 1a vs non type 1a) is initially unsure of classification, with a type 1a probability of around 0.5.
-The probability then decreases slightly, but rapidly increases near the peak of the light curve.
-The classifier has high confidence the supernovae is of type 1a at around 60 days, and the final probability is excess of $99.5\%$.
-This method could therefore be useful for real time monitoring in  surveys.
+As well as finding competitive results for the final metrics, we have shown that it is possible to give fast, early evaluation of supernovae type using pre-trained models.
+This is possible since the light curve can be fed to the model directly without needing any feature extraction.
 
@@
 Further work could also be performed to optimise the early detection probability of the network.
-Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is pre-trained on simulations or existing data from other surveys, then the weights of the network are fine-tuned on the new, smaller dataset.
+Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is pre-trained on simulations or existing data from other surveys then the weights of the network are fine-tuned on the new, smaller dataset.
+The simulated SPCC data used in this work are based on the DES instrument, so it would be interesting to apply transfer learning to real DES data.
 

### 2016-06-23 paper/sn.tex
arXiv ready
Paper

@@
 \centering
-\includegraphics[width=88mm, angle=0]{network.pdf}
+\includegraphics[width=88mm, angle=0]{1.pdf}
 \caption{\label{fig:network} Bidirectional recurrent neural network for sequence classification.
@@
 RNNs have several properties which makes them suitable for sequential information.
-The inputs to the network are flexible, and they are able to recognise patterns with noisy data (for example the context of a word in a sentence relative to others can vary).
+The inputs to the network are flexible, and they are able to recognise patterns with noisy data (for example the context of a word in a sentence relative to others can vary, or a time stream can contain instrument noise).
 
 The main problem with vanilla RNNs  is that they are unable to store long-term information, so inputs at the end of a sequence have no knowledge of inputs at the start.
+This is a problem if the data has long-term correlations.
 Several types of RNNs have been proposed to solve this problem, including {\em Long Short-Term Memory} (LSTM)~\cite{LSTM} and {\em Gated Recurrent Unit} (GRU)~\cite{2014arXiv1412.3555C} units.
@@
 In this case the inputs, which are vectors at each sequential step, are connected to two hidden RNN layers, which can be either vanilla RNN or memory units.
-Each hidden layer contains a number of hidden units, and in each layer information flows either forwards or backwards,  but no information passes between the two directions.
+Each hidden layer contains a number of hidden units (capable of storing information), and in each layer information flows either forwards or backwards,  but no information passes between the two directions.
 Several hidden layers can be stacked to form {\em deep} neural networks.
@@
 
- In the network, each neuron is connected to another by a weight matrix, and the optimal weights are found by back propagating the errors from a {\em loss function} of the output layer.
+ In the network, each neuron is connected to another by a weight matrix, and the optimal weights are found by back-propagating the errors from a {\em loss function} of the output layer.
 For classification problems, this is typically the categorical cross-entropy between predictions and targets, defined as
@@
 where $i,j$ run over the class labels, $t_{i,j}$ are the targets for each class (either 0 or 1) and $p_{i,j}$ are the predicted probabilities.
+Back-propogation essentially takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.
 
@@
 We make 5 random augmentations of all missing data, thereby increasing the size of the dataset by a factor of 5.
-We can test the importance of augmentation by training each augmentation separately and  comparing the change in accuracy.
-We find that augmentation gives a percent level improvement in accuracy.
+We can test the importance of this by training each augmentation separately and  comparing the change in accuracy, which we find is $\sim 1\%$.
+Training with multiple augmentations at once gives the best performance, since the network will learn  to ignore random filled values.
 
@@
 \centering
-\includegraphics[width=88mm, angle=0]{DES_SN551675.pdf}
+\includegraphics[width=88mm, angle=0]{2.pdf}
 \caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g, r, i, z$ bands for SN ID 551675 (a type Ia) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}.
@@
 Each dataset is split into a training and test set, with the training set containing a spectroscopically confirmed supernovae type and redshift.
-It is important that augmented data with the same supernovae ID go into either the training {\em or} test set, not a mix of both.
+It is important that augmented data with the same supernovae ID go into either the training {\em or} test set, not a mix of both, otherwise the training and test data will not be independent.
 The original SPCC data consisted of 1103 training samples.
@@
 \centering
-\includegraphics[width=88mm, angle=0]{lstm_dropout_05_loss.pdf}
+\includegraphics[width=88mm, angle=0]{3.pdf}
 \caption{\label{fig:loss} (Left) Training loss (green) vs test loss (blue) for a unidirectional 2 layer  LSTM network with 16 hidden units in each layer.
@@
 
-We also test the same model using a training fraction of 0.25, around 5000 supernovae, close to the lower  end of the number likely to be followed up for the LSST.
-Again, after 5 randomised runs and training for 200 epochs the accuracy was \comment{$***\pm***$}, the AUC was \comment{$***\pm***$} and \comment{$F_1 =***\pm***$}.
-This suggests that the classifer metrics scale with the data \comment{$\propto***$}.
+We also test the same model using a training fraction of 0.25 (around 5000 supernovae), closer to the lower  end of the number likely to be followed up for the LSST.
+Again, after 5 randomised runs and training for 200 epochs we obtain an accuracy of $93.3 \pm 0.4\%$, an AUC of $0.978 \pm 0.002$ and $F_1 = 0.57 \pm 0.02$.
+The corresponding type Ia purity and completeness are 85\% and 87\% respectively.
+The $F_1$ metric, for example, has degraded by $\sim 10\%$ for a reduction in data of $50\%$.
 
@@
 The loss in the latter case would be lower however, and represents increased confidence of the network in its predictions.
-We therefore reject models with over-fitting and an increasing cross-entropy loss, at the expense of metrics such as $F_1$, and decrease the model complexity.
+We therefore reject models with severe over-fitting and an increasing cross-entropy loss, at the expense of metrics such as $F_1$, and decrease the model complexity.
 
@@
 Further work could also be performed to optimise the early detection probability of the network.
-Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is pre-trained on simulations or existing data from other surveys then the weights of the network are fine-tuned on the new, smaller dataset.
+Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is pre-trained on simulations or existing data from other surveys, then the weights of the network are fine-tuned on the new, smaller dataset.
 The simulated SPCC data used in this work are based on the DES instrument, so it would be interesting to apply transfer learning to real DES data.

### 2016-10-31 paper/sn.tex
LICENSE added
Rewording paper

@@
 Although we have trained the network on the cross-entropy loss  and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-z), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.65).
-Recurrent neural networks  therefore compare very well with a larger training  set.
+Recurrent neural networks  therefore compare well with other methods when a larger training  set is available.
 The performance isn't  quite as good with a smaller training set, possibly due to the network having to learn from no prior information about (noisy) light curves.
-This is also seen in~\cite{Lochner:2016hbn}, which obtains improved metrics for only a small training fraction with a combination of feature extraction and machine learning.
+The current state of the art for a small training set  ($\sim10^3$ supernovae) comes from a combination of SALT2 (Spectral Adaptive Light curve Template 2) template fits and boosted decision trees~\cite{Lochner:2016hbn}.
+It would be interesting to check how how deep learning compares to this with a larger training set.
+
 As well as finding competitive results for the final metrics, we have shown that it is possible to give fast, early evaluation of supernovae type using pre-trained models.

### 2017-01-25 paper/sn_submission_v2/Referee reply.txt
Early-epoch added

To whom it may concern,

We kindly thank the referee for their comments on "Deep recurrent neural networks for supernovae classification". We have addressed each of the points made and feel that this has improved the paper. Laid out below are each of the comments followed by a description of the edits we have added to the paper.

(1) The authors claim that their results are comparable to those from the
challenge itself (several groups had participated in the competition), but they
do not show the comparison.

We have greatly expanded the table to include a range of our results for differing training sample sizes as well as including the results from Karpenka and Newling. It is very difficult to compare our work directly to the other SPCC results since we use (and depend on using) the post-SPCC data. The comparisons are now mentioned in the results section rather than the conclusions.

(2) The original competition was in two parts. The second part was where the
competitors were supposed to use only the first n points to attempt the
classification. Not a single group tried that. I can not require the authors to
attempt that, leave alone succeed. That is a hard problem. But given the new
architectures, that test may be within reach and it will be useful to perhaps
see a discussion. Thats just a suggestion.

We have edited our code to attempt this challenge. We did much better than we expected considering the difficulty of the problem. The network learns really well about the limited initial light-curve data, much better than if we feed only the inital part of the light-curve into a network trained on full light-curves. This is because, when training on the full light-curves the weights are influenced by the later parts of the sequence. This is not true when only using the initial part of the light-curve so the network can learn much better about what happens at the start of the light-curve without worrying about the later part of the data. We have added the results to the last section of the table and mentioned the results in the text.

(3) The authors say that their model could be used with real-data. There are a
very large number of light-curves from many different sky-surveys out there
(though not many in multiple filters). Actually using some of those and
showing that the method actually works will go a long way in convincing anyone.

This is quite a large amount of work which we are currently working on for a future publication. We have mentioned this in the conclusions

(4) Light-curves are used in the input layer. Given that most light-curves have
irregularly spaced points (as well as being heteroskedastic) it will be useful to see
a discussion about if that needs to be taken into consideration for the
construction of the network. For example would it matter if the observations
are bunched up versus well-separated. In general, a better and more detailed 
description of the working.

The points of the light-curves are irregularly spaced. Since this the only available form the data can take we cannot improve on the spacing of the points. Due to the black box nature of recurrent neural networks we cannot be certain whether position in sequence has as great an effect as the value of the time in the sequence, i.e. is the i^th entry of the sequence important or is it the time in that entry. As each sequence needs to be the same length but different light-curves have different sequence lengths shorter sequences are padded and then masked. This could be done, padding all the light-curves to the same sequence length, but then reorganised to put all the similar times at similar sequence entries. We have tested this and found that it is slightly worse with the reorganisation technique. We judge from this that the optimal way  to train the network is to use the light-curves directly and let the network learn about the clustering of points from the values of the times. We can't know for sure, but the network may even learn that there are hidden connections between the clustering of observation times and supernovae type. We have added a paragraph noting this in the text.

(5) Over-fitting, as the authors are aware, can indeed be an issue. It will be
extremely useful if they can isolate the features as seen by the network on
which it discriminates between different types. References to deconvnets 
(used for convnets, not RNNs e.g. arxiv 133.2901) may reveal pieces of code 
that could allow the authors to do so (after some adaption).

It is not easy to demonstrate how to visualise recurrent neural networks (there are few good papers on the topic, mostly on using it in language https://arxiv.org/abs/1506.02078) but these require whole papers devoted to the visualisation. The reason they can't be easily visualised is that the outputs of hidden layers are abstract entities which cannot be mapped back to features of the light-curve. This is different from convolutional neural networks where each output from the hidden convolutional layers are extractions of features from the larger image and is therefore directly related to the original image content.

Finally a few more references to RNNs in the introduction will be useful for
the readers.

We have added a reference to Medsker and Jain's book on recurrent neural networks which is very comprehensive and contains all the information needed to understand RNNs.

Do the authors plan to release the model along with the settled weights?

The code is public so people can train as they wish. We think this is probably more useful to people than having the model and the saved weights.

### 2017-01-25 paper/sn_submission_v3/Referee reply.txt
Early-epoch added

(1) The work as presented is in almost acceptable form. The part about using just the
first few points is powerful and will be welcomed by many. If using the entire
light-curve during testing provides somewhat worse results, then perhaps there
is a way to improve the model even further.

There is possibly a way to improve the model further, but this is not certain. One of our quarms about trying to improve the model further is that the way RNNs work mean that including the information from after the early epoch challenge skews the overall weights of the network meaning that the early epoch weights are less well known compared to training using only the early epoch challenge data.

(2) Many machine learning methods seem like black-boxes and keeping them that way
does not help dispel myths about them from the eyes of the uninitiated. When
describing example data, can some stats be provided about the amount of
chunking in light-curves? For example the range of time-span for the
light-curves, the number of chunks they are typically divided in (a box-plot
could be used to describe median, show outliers etc.) Or point to some
specifics from the SPCC figures (table X, Fig Y etc.)

(3) *ANY* discussion about error-bars will be helpful in letting readers decide if
they would want to apply the technique. [See comment below about 73+/-2%]. But
thats just a suggestion and not a requirement.

The errors in the results are due to running the training 5 times with different random selections of training data (and random augmentations of data). This is mentioned in the results section already.

(4) Did the authors try removing known points, then apply their chunking and random
filling and see if they get consistent results?

To do...

(5) The discussion about padding is unclear. if, in the end, the padding does not
get used, may be the discussion should be removed to an appendix or a footnote
(but do not remove it as it is part of the flow). Perhaps an example will help?

The padding is used to create the vectors of the correct length. This padding is then ignored by the network using a technique called masking. This means the padding has no effect on the results, it just maintains the correct length vectors. This is already explicitly described in the text.

(6a) In Table 2 please include the 0.052% even if the results are bad.

They are already in the table in every case - this a train fraction of 0.052, not 0.052%.

(6b) Are the early-epoch results for SN1a or 123? Can the results be provided for both sets?

The early-epoch challenge is for SN1a only and not 123. We have given that a go and added the results to the table.

(6c) In the middle section, in the last 2 lines, Completeness seems to rise when Host
information is dropped. Thats anomalous. If you understand why, a comment will
be useful. Otherwise those are the kind of error-bars that become meaningful:
are both 73+/-2%? You can still fit the table on a single page by abbreviating
contents in the first column.

This was a statistical fluctuation. We have added the errors on the purity and completeness. As it can be seen the errors on both of the bottom lines of the second section are ~4%, approximately the same as the difference between the means.

(7) It is understandable that your methods are not comparable with various methods
used during SPCC, but perhaps its still worth mentioning the best numbers
there?

To do...

(8) The early-epoch result is surprising. May be it should be emphasized more? It
will be useful to many people.

To do..

(9) The following is a loose sentence: "Moreover, we have ...trained on the whole
light-curve". Especially so since you do not normalize time to the peak value so the
starting point could be arbitrary (or is your padding mitigating exactly that?
if so, clarify a little). Otherwise quantify (at least verbally) what you mean
by the above sentence.

We mean that training using the data from the early epoch challenge gives a better early-time predictor than training using the whole light-curve. We have rewritten the sentence to make it clearer.

(10) You mention that perhaps even raw images could be used etc., but you have not
even discussed the effects of different filters etc. What happens if you use
just one or two filters, for instance? So, either go into more details, or
exclude pure speculation.

This isn't pure speculation, this is a consideration for further work. It is possible to data mine images and pipe this into an RNN. We could do this for a single filter, we could do this for many filters. If one filter was missing then the input could be augmented as done here and the other raw images convolved to extract features. This is identical to what is done here, but with a huge amount more information.

(11) Section 4: Over-fitting can also show up as some inputs being learnt too well
(and thus its not the noise that gets fitted)

What is different between differences in inputs and noise? They are the same as far as we can tell

(12) You mention in the comments that the code is public. Please refer to it so that
it can be used by others.

This is already mentioned, the link is on the front page.

(13) Shorten the y-axes in figure 3 (keeping it same for both parts) to reduce
unused white-space.

Done.

minor comments:
(1) references need to be delineated better - many seem to be inline now and don't feel right.

We are not sure how it's supposed to feel?

(2) inputted -> input

We have changed this

### 2017-01-25 paper/sn_submission/ms.tex
Early-epoch added

--- 2016-10-31 paper/sn.tex
@@
 %\documentclass{article}
-\documentclass[nofootinbib,amsmath,amssymb,10pt,eqsecnum, twocolumn]{revtex4-1}
+%\documentclass[nofootinbib,amsmath,amssymb,10pt,eqsecnum, twocolumn]{aastex61}
+\documentclass[twocolumn]{aastex61}
 \usepackage{graphicx, amsmath, amsmath, amssymb}
 \usepackage[caption=false]{subfig}
-\usepackage[breaklinks, colorlinks, citecolor=blue]{hyperref}
+%\usepackage[breaklinks, colorlinks, citecolor=blue]{hyperref}
 %\usepackage[maxbibnames=4]{biblatex}
-\usepackage{color}
+%\usepackage{color}
 \usepackage{float}
 \usepackage{natbib,hyperref,ifthen}
+\usepackage{subfigure}
 %------------ May need to comment out this -----------
@@
 
-\bibliographystyle{arxiv}
-
 \def\mnras{MNRAS}
@@
 
-\title{Deep Recurrent Neural Networks for Supernovae Classification}
+\title{Deep recurrent neural networks for supernovae classification}
 
@@
 Using the Supernovae Photometric Classification Challenge (SPCC) data, we find that deep networks are capable of learning about light curves, however the performance of the network is highly sensitive to the amount of training data.
-For a training size of 50\% of the representational SPCC dataset (around $10^4$ supernovae) we obtain a type Ia vs non type Ia classification accuracy of 94.8\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$.
+For a training size of 50\% of the representational SPCC dataset (around $10^4$ supernovae) we obtain a type-Ia vs.
+non-type-Ia classification accuracy of 94.8\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$.
 We also apply a pre-trained model to obtain classification probabilities as a function of time, and show it can give early indications of supernovae type.
@@
 
-\maketitle
+%\maketitle
 
@@
 The amount of data produced lends itself to new analysis methods which can learn abstract representations of complex data.
-Deep learning is a powerful method for gaining multiple levels of abstraction, and has recently produced state-of-the-art results in tasks such as image classification and natural language processing.
-There are several types of deep learning architectures, such as convolutional neural networks, deep belief networks and recurrent neural networks (see~\cite{0483bd9444a348c8b59d54a190839ec9} for an excellent overview of deep learning and refs.
+Deep learning is a powerful method for gaining multiple levels of abstraction, and has recently produced state-of-the-art results in tasks such as image classification and natural language processing (see~\cite{0483bd9444a348c8b59d54a190839ec9} for an excellent overview of deep learning and refs.
 within for more details).
@@
 \centering
-\includegraphics[width=88mm, angle=0]{1.pdf}
+\includegraphics[width=88mm, angle=0]{f1.pdf}
 \caption{\label{fig:network} Bidirectional recurrent neural network for sequence classification.
@@
 This is a problem if the data has long-term correlations.
-Several types of RNNs have been proposed to solve this problem, including {\em Long Short-Term Memory} (LSTM)~\cite{LSTM} and {\em Gated Recurrent Unit} (GRU)~\cite{2014arXiv1412.3555C} units.
+Several types of RNNs have been proposed to solve this problem, including {\em Long Short-Term Memory} (LSTM) units~\cite{LSTM} and {\em Gated Recurrent Units} (GRU)~\cite{2014arXiv1412.3555C}.
 These are similar in concept, in that information is able to flow through the network via a gating mechanism.
-One further problem with RNNs is that information is only able to flow in one direction.
+Another problem with RNNs is that information can only flow in one direction.
 In {\em bidirectional} RNNs information is able to pass both forwards and backwards.
-Bidirectional LSTM networks have been shown to be particularly powerful for supervised sequence labelling, where sequential data is accompanied by a set of discrete labels.
+Bidirectional LSTM networks have been shown to be particularly powerful where sequential data is accompanied by a set of discrete labels.
 
 The architecture of a typical bidirectional RNN for sequence labelling is shown in Fig.~\ref{fig:network}, where the squares represent {\rm neurons}.
-In this case the inputs, which are vectors at each sequential step, are connected to two hidden RNN layers, which can be either vanilla RNN or memory units.
+In this case the inputs, which are vectors at each sequential step, are connected to two hidden RNN layers, either vanilla RNN or memory units.
 Each hidden layer contains a number of hidden units (capable of storing information), and in each layer information flows either forwards or backwards,  but no information passes between the two directions.
@@
 The output from the final set of hidden layers in each direction is merged at each sequential step, and mean pooled (averaged) over all steps to obtain a consensus view of the network\footnote{We find that obtaining a consensus view improves the performance of the network.}.
-Finally, the mean output is fed to a {\em softmax} layer, which takes an input vector ${\bf z}$ and returns normalised, exponentiated outputs for each class label $i$, $\exp(z_i) / \sum_{i} \exp(z_i)$, i.e.
+Finally, the mean output is fed to a {\em softmax} layer, taking an input vector ${\bf z}$ and returning normalised, exponentiated outputs for each class label $i$, $\exp(z_i) / \sum_{i} \exp(z_i)$, i.e.
 a vector of probabilities.
 
- In the network, each neuron is connected to another by a weight matrix, and the optimal weights are found by back-propagating the errors from a {\em loss function} of the output layer.
+ Each neuron is connected to another by a weight matrix, and the optimal weights are found by back-propagating the errors from a {\em loss function} of the output layer.
 For classification problems, this is typically the categorical cross-entropy between predictions and targets, defined as
@@
 where $i,j$ run over the class labels, $t_{i,j}$ are the targets for each class (either 0 or 1) and $p_{i,j}$ are the predicted probabilities.
-Back-propogation essentially takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.
+Back-propogation takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.
 
@@
 
-In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, which consists of a set of 21319 simulated supernove light curves.
-Each supernovae sample consists of a time series of flux measurements, with errors, in the $g, r, i, z$ bands (one band for each timestep), along with the position on the sky and dust extinction.
+In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, consisting of 21319 simulated supernove light curves.
+Each supernovae sample consists of a time series of flux measurements, with errors, in the $g,r,i,z$ bands (one band for each timestep), along with the position on the sky and dust extinction.
 An example set of light curves is shown in Fig.~\ref{fig:lightcurve}.
 
-Due to the format of the input data, we first do a small amount of data processing to obtain values of the $g, r, i, z$ fluxes and errors at each sequential step.
-First, we assume the time sequence begins at day 0 for each supernovae, rather than counting days forwards and backwards from the maxima of the light curve.
-Next, for observations less than $\sim1$ hour apart, we group the $g, r, i, z$ values into a single vector, ensuring there is at most one filter type in each group.
-If there is more than one filter type, we further subdivide the group using a finer time interval.
-We take the time of the group to be the mean time of its individual observations, which is reasonable as the time intervals are small compared to the characteristic time of the light curve.
+Due to the format of the input data, we first do a small amount of data processing to obtain values of the $g,r,i,z$ fluxes and errors at each sequential step.
+We assume the time sequence begins at day 0 for each supernovae, rather than counting days forwards and backwards from the maxima of the light curve.
+For observations less than $\sim1$ hour apart, we group the $g,r,i,z$ values into a single vector, ensuring there is at most one filter-type in each group.
+If there is more than one filter-type, we further subdivide the group using a finer time interval.
+The group time is the mean of the times of each observation, which is reasonable as the time intervals are small compared to the characteristic time of the light curve.
 
- Observations are then of the form in Tab.~\ref{tab:augment}, where any missing values are denoted by a dash.
+ Observations are of the form in Table~\ref{tab:augment}, where any missing values are denoted by a dash.
 In order to impute the missing value of $i$, we use {\em data augmentation} and randomly select a value between $i_1$ or $i_3$.
-We make 5 random augmentations of all missing data, thereby increasing the size of the dataset by a factor of 5.
+We make 5 random augmentations of all missing data, thereby increasing the size of the dataset fivefold.
 We can test the importance of this by training each augmentation separately and comparing the change in accuracy, which we find is $\sim 1\%$.
-Training with multiple augmentations at once gives the best performance, since the network will learn  to ignore random filled values.
+Training with multiple augmentations at once gives the best performance since the network learns to ignore random-filled values.
 
@@
 \begin{center}
-\begin{tabular}{|c|c|c|c|c|} \hline  
- Time & g & r & i & z \\ \hline \hline 
-$ t_1$  & $g_1$ & $r_1$ &$ i_1$ & $z_1$ \\ \hline 
-$ t_2$  & $g_2$ & $r_2$ &$  - $ & $z_2$ \\ \hline 
-$ t_3$  & $g_3$ & $r_3$ &$ i_3$ & $z_3$ \\ \hline 
+\begin{tabular}{|c|c|c|c|c|}\hline  
+ Time&g&r&i&z\\\hline\hline 
+$t_1$&$g_1$&$r_1$&$i_1$&$z_1$\\\hline 
+$t_2$&$g_2$&$r_2$&$-$&$z_2$\\\hline 
+$t_3$&$g_3$&$r_3$&$i_3$& $z_3$\\\hline 
  \end{tabular}
@@
 \centering
-\includegraphics[width=88mm, angle=0]{2.pdf}
-\caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g, r, i, z$ bands for SN ID 551675 (a type Ia) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}.
-The data has been processed using augmentation so there is a $g, r, i, z$ value at each sequential step.
-(Bottom) Type Ia probability as a function of time from a 2 layer LSTM model, trained with around $10^4$ supernovae and SN 551675 excluded.
-The final probability gives $99.5\%$ confidence that the supernovae is of type Ia.
+\includegraphics[width=88mm, angle=0]{f2.pdf}
+\caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g,r,i,z$ bands for SN ID 551675 (a type-Ia) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}.
+The data has been processed using augmentation so there is a $g,r,i,z$ value at each sequential step.
+(Bottom) Type-Ia probability as a function of time from a 2 layer LSTM model, trained with around $10^4$ supernovae and SN 551675 excluded.
+The final probability gives $99.5\%$ confidence that the supernovae is of type-Ia.
  }
@@
 Each dataset is split into a training and test set, with the training set containing a spectroscopically confirmed supernovae type and redshift.
-It is important that augmented data with the same supernovae ID go into either the training {\em or} test set, not a mix of both, otherwise the training and test data will not be independent.
+It is important that augmented data with the same supernovae ID go into either the training {\em or} test set otherwise they will not be independent.
 The original SPCC data consisted of 1103 training samples.
@@
 
-The input vector to each sequential step then consists of: the time in days since the first observation; the flux in each of the 4 bands; the flux errors in each of the 4 bands; the RA and Dec; the dust extinction; and the host photo-z if relevant.
+The input vector to each sequential step consists of: time in days since the first observation; flux in each of the 4 bands; flux errors in each of the 4 bands; RA and Dec; dust extinction; and host photo-z if relevant.
 Whilst we do not expect some of these variables to impact the classifier accuracy, we do not attempt any feature engineering and leave it to the network to decide if they are relevant.
 
-RNNs typically perform better with more training data, so for training purposes we use the SPCC test set with answer keys (which is a non-biased representational dataset\footnote{The original SPCC training set was non-representational.}), and select a random fraction to act as the training set.
+RNNs typically perform better with more training data, so we train using the SPCC test set with answer keys (which is a non-biased representational dataset\footnote{The original SPCC training set was non-representational.}), and select a random fraction to act as the training set.
 We consider 1103 supernovae (a training fraction of 0.052), the same size as the original challenge, and fractions of 0.25 and 0.5 (around 5000 and $10^4$ supernovae respectively), nearly an order of magnitude larger, and closer to the number likely to be followed up for the LSST.
@@
 The goal of the classifier is to determine the supernovae type in the test set.
-We consider two problems, the first being to categorise two classes (type Ia vs non type Ia), and the second to categorise three classes (supernovae types 1, 2 and 3).
+We consider two problems, (1) to categorise two classes (type-Ia vs.
+non-type-Ia), and (2) to categorise three classes (supernovae types-1, -2 and -3).
 We denote these as  `SN1a' and  `123' respectively.
 
-We use several metrics to assess the classifier.
+Several metrics are used to assess the classifier.
 The simplest is the accuracy, defined as the ratio between the number of correct predictions and total number of predictions.
@@
 Next are a variety of metrics coming from the {\em confusion matrix} of predictions.
-For binary classification problems, the confusion matrix splits predictions into true positives (TP), false positives (FP),  false negatives (FN), and true negatives (TN).
-We consider the purity (also known as precision) and completeness (also known as recall) of the classifier.
+For binary classification problems, the confusion matrix splits predictions into true positives~(TP), false positives~(FP), false negatives~(FN), and true negatives~(TN).
+We consider the purity and completeness of the classifier.
 These are defined as
@@
 \end{equation}
-We evaluate these for each class separately vs `the rest' (e.g. type Ia vs non type Ia). The SPCC also defined the $F_1$ figure-of-merit for the SN1a classification problem. This is 
+We evaluate these for each class separately vs. `the rest' (e.g. type-Ia vs. non-type-Ia). The SPCC also defined the $F_1$ figure-of-merit for the SN1a classification problem. This is 
 \begin{equation}
@@
 \end{equation}
-so false positives (incorrectly classifying a non type Ia supernovae as a type Ia) are penalised more heavily.
+so incorrectly classifying a non-type-Ia supernovae as a type-Ia is penalised more heavily.
 
-Finally, we also calculate  the Area Under the Curve (AUC).
-The AUC is the area under the curve of the TP rate vs FP rate, as the threshold probability for classification is increased from 0 to 1.
+Finally, we calculate the Area-Under-the-Curve (AUC).
+The AUC is the area under the curve of the TP rate vs.
+FP rate, as the threshold probability for classification is increased from 0 to 1.
 A perfect classifier has an AUC of 1, and a random classifier 0.5.
-For multi-class problems, we calculate the AUC for each class vs the rest, and take an unweighted average to give the final AUC score.
+For multi-class problems, we calculate the AUC for each class vs.
+the rest, and take an unweighted average to give the final AUC score.
 
@@
 We also consider unidirectional and bidirectional networks.
-For the unidirectional network, we fix the direction to be forwards.
-For the bidirectional network, we fix the number of hidden units in each RNN layer to be equal for the forward and backward directions.
+For unidirectional networks we fix the direction to be forwards.
+For bidirectional networks, the number of hidden units in each RNN layer is equal in the forward and backward directions.
 
@@
 We denote the number of hidden units in a network with a  single stack by $[h_1]$, and the number of hidden layers in a two stack model by $[h_1, h_2]$.
-We vary the number of hidden units, testing h=[4], [8], [16], [32], [4,4], [8,8], [16,16] and [32,32].
+We vary the number of hidden units, testing $h=[4],[8],[16],[32],[4,4],[8,8],[16,16]$ and~$[32,32]$.
 We do not go beyond a stack of two layers due to the limited size of the dataset.
@@
 For each network we perform 5 randomised runs over the training data to obtain the classifier metrics.
-We define the loss function as the categorical cross-entropy between the predictions and test data.
-The network weights were trained using back-propogation with the `Adam' updater~\cite{2014arXiv1412.6980K}.
-Mini-batches of size 10\footnote{If training with a GPU larger mini-batches are recommended to make use of the GPU cores.} were used throughout, and each model was trained over approximately 200 {\em epochs}, where each epoch is a full pass over the training data.
+The loss function is the categorical cross-entropy between the predictions and test data.
+The network weights ware trained using back-propagation with the {\ttfamily Adam} updater~\cite{2014arXiv1412.6980K}.
+Mini-batches containing 10 samples\footnote{If training with a GPU larger mini-batches are recommended to make use of the GPU cores.} were used throughout, and each model was trained for 200 {\em epochs}, where each epoch is a full pass over the training data.
 
@@
 \centering
-\includegraphics[width=88mm, angle=0]{3.pdf}
-\caption{\label{fig:loss} (Left) Training loss (green) vs test loss (blue) for a unidirectional 2 layer  LSTM network with 16 hidden units in each layer.
-(Right) Training accuracy (green) vs test accuracy (blue) for the same network.
+\includegraphics[width=88mm, angle=0]{f3.pdf}
+\caption{\label{fig:loss} (Left) Training loss (green) vs.
+test loss (blue) for a unidirectional 2 layer LSTM network with 16 hidden units in each layer.
+(Right) Training accuracy (green) vs.
+test accuracy (blue) for the same network.
 }
@@
 A dataset of 21319 is relatively small by deep learning standards.
-Furthermore, the `feature space' of supernovae light curves is significantly smaller than, say using RNNs to learn about language.
+Furthermore, the `feature space' of supernovae light curves is significantly smaller than, say, using RNNs to learn about language.
 We therefore need to be careful about {\em over-fitting}.
@@
 If the loss of training data continues to decrease, but the loss of the test data increases, this is a sure sign of over-fitting.
-On the other hand, if no sign of over-fitting is observed, the network is not usually complex enough to fully learn the relationship between inputs and outputs (called {\em under-fitting}).
+If no sign of over-fitting is observed, the network is not usually complex enough to fully learn the relationship between inputs and outputs (called {\em under-fitting}).
 
-For a training fraction of 0.5, we found the best architecture was a deep 2 layer network with unidirectional LSTM units.
+For a training fraction of 0.5, we found the best architecture was a deep 2-layer network with unidirectional LSTM units.
 Bidirectional units did not significantly improve the test accuracy and made the network more difficult to train.
-There was a marked improvement in test accuracy using 16 hidden units in each layer rather than 8, but too much over-fitting occured using 32 hidden units.
-Over-fitting was still an issue for 16 hidden units,  but we make use of a technique called {\em dropout}~\cite{JMLR:v15:srivastava14a} to regularize this.
-Dropout  sets a random fraction of connections to 0 at each update during training {\em only}, which prevents the units from adapting too much.
+There was a marked improvement in test accuracy using 16 hidden units in each layer rather than 8, but too much over-fitting occurred using 32 hidden units.
+Over-fitting was still an issue for 16 hidden units,  but a technique called {\em dropout}~\cite{JMLR:v15:srivastava14a} could regularise this.
+Dropout sets a random fraction of connections to 0 at each update during training {\em only}, preventing the units from adapting too much.
 We apply dropout only to non-recurrent connections after each hidden layer.
 
-In Fig.~\ref{fig:loss} we show the training and tests losses for such a network, with a dropout of $0.5$, applied to type Ia vs non type Ia classification with host galaxy photo-z information.
-Without dropout the training loss would continue to fall and the test loss would rise.
-For 5 randomised runs, after training for  200 epochs, we obtain a classification accuracy of $94.8 \pm 0.2$\%, an AUC of $0.986 \pm 0.001$ and $F_1 = 0.64 \pm 0.01$.
-The corresponding type Ia purity and completeness are 87\% and 92\% respectively.
-A summary of results for this network can be found in Tabs.~\ref{tab:05_SN1a} and~\ref{tab:05_123}.
-It can be seen that the inclusion of host galaxy photo-z does not significantly improve the classifier performance.
+In Fig.~\ref{fig:loss} we show the training and tests losses for such a network, with a dropout of $0.5$, applied to type-Ia vs.
+non-type-Ia classification with host galaxy photo-z information.
+Without dropout the training loss continues to fall and the test loss rises.
+For 5 randomised runs, training for 200 epochs, we obtain a classification accuracy of $94.8 \pm 0.2$\%, AUC of $0.986 \pm 0.001$ and $F_1 = 0.64 \pm 0.01$.
+The corresponding type-Ia purity and completeness are 87\% and 92\% respectively.
+A summary of results for this network can be found in Tables~\ref{tab:05_SN1a} and~\ref{tab:05_123}.
+The inclusion of host galaxy photo-z does not significantly improve the classifier performance.
 
 \begin{table}[t!]
+\begin{subtable}
+\centering
 \begin{tabular}{|c | c | c | c | c |} \hline 
@@
 \end{tabular}
-\caption{\label{tab:05_SN1a} Summary of results for  type Ia vs non type Ia classification with a training fraction of 0.5.
-The network used is a 2 layer LSTM, each with 16 hidden units, and a dropout of 0.5.}
-\end{table}
-
-\begin{table}[t!]
+\caption{\label{tab:05_SN1a} Summary of results for  type-Ia vs.
+non-type-Ia classification with a training fraction of 0.5.
+The network is a 2-layer LSTM, each with 16 hidden units and dropout of 0.5.}
+\end{subtable}%
+\begin{subtable}
+\centering
 \begin{tabular}{|c | c | c | c | c |} \hline 
@@
 \end{tabular}
-\caption{\label{tab:05_123} Summary of results for three class categorisation (supernovae types 1, 2 and 3). The model is the same as in Tab.~\ref{tab:05_SN1a}}
+\caption{\label{tab:05_123} Summary of results for three class categorisation (supernovae types-1, -2 and -3). The model is the same as in Table~\ref{tab:05_SN1a}}
+\end{subtable}
 \end{table}
 
-One advantage of our approach is that light curve data can be directly inputted to a {\em pre-trained} model to give very fast evaluation ($<1\,$s) of supernovae type.
-In the lower panel of Fig.~\ref{fig:lightcurve} we input the light curve, as a function of time, of a type Ia supernovae (excluded from training) to the pre-trained 2 layer LSTM model discussed above.
-The classifier (type Ia vs non type Ia) is initially unsure of classification, with a type Ia probability of around 0.5.
+One advantage of our approach is that light curve data can be directly inputted to a {\em pre-trained} model to give very fast evaluation ($<1$s) of supernovae type.
+In the lower panel of Fig.~\ref{fig:lightcurve} we input the light curve, as a function of time, of a type-Ia supernovae (excluded from training) to the pre-trained 2-layer LSTM model discussed above.
+The classifier (type-Ia vs.
+non-type-Ia) is initially unsure of classification, with a type-Ia probability of around 0.5.
 The probability then decreases slightly, but rapidly increases near the peak of the light curve.
-The classifier has high confidence the supernovae is of type Ia at around 60 days, and the final probability is excess of $99.5\%$.
+The classifier has high confidence the supernovae is of type-Ia at around 60 days, and the final probability is excess of $99.5\%$.
 This method could therefore be useful to give early indication of supernovae type in surveys.
@@
 We also test the same model using a training fraction of 0.25 (around 5000 supernovae), closer to the lower end of the number likely to be followed up for the LSST.
-Again, after 5 randomised runs and training for 200 epochs we obtain an accuracy of $93.3 \pm 0.4\%$, an AUC of $0.978 \pm 0.002$ and $F_1 = 0.57 \pm 0.02$.
-The corresponding type Ia purity and completeness are 85\% and 87\% respectively.
-The $F_1$ metric, for example, has degraded by $\sim 10\%$ for a reduction in data of $50\%$.
+After 5 randomised runs and training for 200 epochs we obtain an accuracy of $93.3 \pm 0.4\%$, AUC of $0.978 \pm 0.002$ and $F_1 = 0.57 \pm 0.02$.
+The corresponding type-Ia purity and completeness are 85\% and 87\% respectively.
+The $F_1$ metric has degraded by $\sim10\%$ for a reduction in data of $50\%$.
 
 For $5.2\%$ of the representative SPCC data, the training dataset is so small that over-fitting is more severe.
-Using the same 2 layer LSTM network with 16 hidden units and a dropout of 0.5 we find a notable increase in the test loss after $\sim 20$ epochs, but the accuracy and other metrics remain relatively constant ($F_1$ values, for example, of 0.35 to 0.4 were obtained).
+Using the same 2-layer LSTM network with 16 hidden units and dropout of 0.5 we find a notable increase in the test loss after $\sim 20$ epochs, but the accuracy and other metrics remain relatively constant ($F_1$ values of 0.35 to 0.4 were obtained).
 The reason for this apparent discrepancy is that the accuracy, say, simply takes the maximum value of the softmax output layer.
@@
 
-For a training fraction of  $5.2\%$ we find a single layer LSTM network, with 4 hidden units, and a dropout of 0.5 satisfies this criteria.
-For 5 randomised runs, after training for  200 epochs, we obtain a classification accuracy of $85.9 \pm 1.2$\%, an AUC of $0.910 \pm 0.012$ and $F_1 = 0.31 \pm 0.03$.
-The corresponding type Ia purity and completeness are 72\% and 66\% respectively.
+For a training fraction of  $5.2\%$ we find a single-layer LSTM network, with 4 hidden units, and dropout of 0.5 satisfies this criteria.
+For 5 randomised runs, training for 200 epochs, we obtain a classification accuracy of $85.9 \pm 1.2$\%, AUC of $0.910 \pm 0.012$ and $F_1 = 0.31 \pm 0.03$.
+The corresponding type-Ia purity and completeness are 72\% and 66\% respectively.
 
@@
 Machine learning methodology has previously been applied to SPCC classification~\cite{Newling:2010bp, Karpenka:2012pm, Lochner:2016hbn}.
-These all employ a two step process, where features are first extracted by some method before classification.
+These employ a two-step process, where features are first extracted by some method before classification.
 In our approach light curves are used directly as inputs to a recurrent neural network, which is able to learn information from the sequence of observations.
 
-Using post SPCC data, we obtain results which are competitive with previous approaches.
+Using post-SPCC data, we obtain results which are competitive with previous approaches.
 Although we have trained the network on the cross-entropy loss and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-z), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.65).
@@
 The performance isn't  quite as good with a smaller training set, possibly due to the network having to learn from no prior information about (noisy) light curves.
-The current state of the art for a small training set  ($\sim10^3$ supernovae) comes from a combination of SALT2 (Spectral Adaptive Light curve Template 2) template fits and boosted decision trees~\cite{Lochner:2016hbn}.
+The current state-of-the-art for a small training set ($\sim10^3$ supernovae) comes from a combination of SALT2 (Spectral Adaptive Light curve Template 2) template fits and boosted decision trees~\cite{Lochner:2016hbn}.
 It would be interesting to check how how deep learning compares to this with a larger training set.
@@
 We appreciate helpful conversations with Steven Bamford, Simon Dye, Mark Sullivan and Michael Wood-Vasey, and Natasha Karpenka, Richard Kessler and Michelle Lochner for help with data acquisition.
-T.
-C.
-is supported by a STFC studentship, and A.
-M.
+T.C.
+is supported by a STFC studentship, and A.M.
 is supported by a Royal Society University Research Fellowship.
 
-\bibliography{bib}
+\begin{thebibliography}{}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+\providecommand{\url}[1]{\href{#1}{#1}}
+
+\bibitem[{{Chung} {et~al.}(2014){Chung}, {Gulcehre}, {Cho}, \&
+  {Bengio}}]{2014arXiv1412.3555C}
+{Chung}, J., {Gulcehre}, C., {Cho}, K., \& {Bengio}, Y. 2014, ArXiv e-prints,
+  arXiv:1412.3555
+
+\bibitem[{Hochreiter \& Schmidhuber(1997)}]{LSTM}
+Hochreiter, S., \& Schmidhuber, J. 1997, Neural Computation, 9(8), 1735
+
+\bibitem[{Karpenka {et~al.}(2013)Karpenka, Feroz, \& Hobson}]{Karpenka:2012pm}
+Karpenka, N.~V., Feroz, F., \& Hobson, M.~P. 2013, Mon. Not. Roy. Astron. Soc.,
+  429, 1278
+
+\bibitem[{Kessler {et~al.}(2010{\natexlab{a}})Kessler, Conley, Jha, \&
+  Kuhlmann}]{Kessler:2010wk}
+Kessler, R., Conley, A., Jha, S., \& Kuhlmann, S. 2010{\natexlab{a}},
+  arXiv:1001.5210
+
+\bibitem[{Kessler {et~al.}(2010{\natexlab{b}})}]{Kessler:2010qj}
+Kessler, R., {et~al.} 2010{\natexlab{b}}, Publ. Astron. Soc. Pac., 122, 1415
+
+\bibitem[{{Kingma} \& {Ba}(2014)}]{2014arXiv1412.6980K}
+{Kingma}, D., \& {Ba}, J. 2014, ArXiv e-prints, arXiv:1412.6980
+
+\bibitem[{Lecun {et~al.}(2015)Lecun, Bengio, \&
+  Hinton}]{0483bd9444a348c8b59d54a190839ec9}
+Lecun, Y., Bengio, Y., \& Hinton, G. 2015, Nature, 521, 436
+
+\bibitem[{Lochner {et~al.}(2016)Lochner, McEwen, Peiris, Lahav, \&
+  Winter}]{Lochner:2016hbn}
+Lochner, M., McEwen, J.~D., Peiris, H.~V., Lahav, O., \& Winter, M.~K. 2016,
+  arXiv:1603.00882
+
+\bibitem[{{LSST Science Collaboration} {et~al.}(2009){LSST Science
+  Collaboration}, {Abell}, {Allison}, {Anderson}, {Andrew}, {Angel}, {Armus},
+  {Arnett}, {Asztalos}, {Axelrod}, \& et~al.}]{2009arXiv0912.0201L}
+{LSST Science Collaboration}, {Abell}, P.~A., {Allison}, J., {et~al.} 2009,
+  ArXiv e-prints, arXiv:0912.0201
+
+\bibitem[{{Matheson} {et~al.}(2013){Matheson}, {Fan}, {Green}, {McConnachie},
+  {Newman}, {Olsen}, {Szkody}, \& {Wood-Vasey}}]{2013arXiv1311.2496M}
+{Matheson}, T., {Fan}, X., {Green}, R., {et~al.} 2013, ArXiv e-prints,
+  arXiv:1311.2496
+
+\bibitem[{Newling {et~al.}(2011)}]{Newling:2010bp}
+Newling, J., {et~al.} 2011, Mon. Not. Roy. Astron. Soc., 414, 1987
+
+\bibitem[{Srivastava {et~al.}(2014)Srivastava, Hinton, Krizhevsky, Sutskever,
+  \& Salakhutdinov}]{JMLR:v15:srivastava14a}
+Srivastava, N., Hinton, G., Krizhevsky, A., Sutskever, I., \& Salakhutdinov, R.
+  2014, Journal of Machine Learning Research, 15, 1929.
+\newblock \url{http://jmlr.org/papers/v15/srivastava14a.html}
+
+\end{thebibliography}
 

### 2017-01-25 paper/sn_submission_v3/ms.tex
Early-epoch added

--- 2017-01-25 paper/sn_submission/ms.tex
@@
 \usepackage{subfigure}
+\usepackage{array}
 %------------ May need to comment out this -----------
@@
 \newcommand\comment[1]{\textcolor{red}{#1}}
+\newcommand{\specialcell}[2][c]{\renewcommand{\arraystretch}{0.8}%
+  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
 
@@
 
-{\em Recurrent neural networks} (RNNs) are a class of artificial neural network  that can learn about sequential data.
+{\em Recurrent neural networks} (RNNs) are a class of artificial neural network  that can learn about sequential data (for an extremely comprehensive overview see~\cite{medsker1999recurrent}).
 They are commonly used for tasks such speech recognition and language translation, but have several possible applications in astronomy  and cosmology for processing temporal or spatial sequential data.
@@
 Observations are of the form in Table~\ref{tab:augment}, where any missing values are denoted by a dash.
-In order to impute the missing value of $i$, we use {\em data augmentation} and randomly select a value between $i_1$ or $i_3$.
+In order to impute the missing value of $i$, we use {\em data augmentation} and randomly select a value between $i_1$ and $i_3$.
 We make 5 random augmentations of all missing data, thereby increasing the size of the dataset fivefold.
@@
 \end{center}
-\label{tab:augment}
-\caption{Data augmentation of missing observations. The missing data is replaced randomly by a value between $i_1$ or $i_3$.}
+\caption{\label{tab:augment} Data augmentation of missing observations. The missing data is replaced randomly by a value between $i_1$ and $i_3$.}
 \end{table}%
@@
 
+The times of the observations in the light-curve are irregularly spaced and whilst this may not be optimal for the network we find that it is better to use the data padded at the end of the sequence than to place observations at similar times in similar sequence positions.
+There may even be hidden connections between the clustering of observation times and supernovae type, although it is hard to test for this.
+
 The goal of the classifier is to determine the supernovae type in the test set.
@@
 We denote these as  `SN1a' and  `123' respectively.
+We also attempt to categorise between type-Ia and non-type-Ia supernovae using only the first six observations with $S/N>4$ and the data taken on the night of the sixth observation as described in~\cite{Kessler:2010wk}.
 
@@
 Without dropout the training loss continues to fall and the test loss rises.
-For 5 randomised runs, training for 200 epochs, we obtain a classification accuracy of $94.8 \pm 0.2$\%, AUC of $0.986 \pm 0.001$ and $F_1 = 0.64 \pm 0.01$.
-The corresponding type-Ia purity and completeness are 87\% and 92\% respectively.
-A summary of results for this network can be found in Tables~\ref{tab:05_SN1a} and~\ref{tab:05_123}.
-The inclusion of host galaxy photo-z does not significantly improve the classifier performance.
+For 5 randomised runs, training for 200 epochs, we obtain a classification accuracy of $94.9 \pm 0.2$\%, AUC of $0.986 \pm 0.001$ and $F_1 = 0.64 \pm 0.01$.
+The corresponding type-Ia purity and completeness are $87.3 \pm 0.8\%$ and $91.4\pm 1.1\%$ respectively.
+A summary of results and comparisons can be found in Table~\ref{tab:05_SN1a}.
+The inclusion of host galaxy photo-$z$ marginally improves the classifier performance.
 
-\begin{table}[t!]
-\begin{subtable}
+\begin{table*}[t!]
 \centering
-\begin{tabular}{|c | c | c | c | c |} \hline 
-AUC & Accuracy & F1 & Host z \\ \hline 
-$ 0.986 \pm  0.001$ &  $  94.8 \pm 0.2\% $ & $  0.64 \pm 0.01  $ & T  \\
-$ 0.981 \pm  0.002$ &  $  93.6 \pm 0.3\% $ & $  0.60 \pm 0.02  $ & F  \\ \hline
+\begin{tabular}{ c  c  c  c  c  c  c  c }\hline\hline
+Method & Training size & AUC & Accuracy (\%) & $F_1$ & Purity (\%) & Completeness (\%) & Host $z$ \\\hline\hline
+A & 10,660 & $0.986 \pm 0.001$ & $94.7 \pm 0.2$ & $0.64 \pm 0.01$ & $87.3 \pm 0.8$ & $91.4 \pm 1.1$ & True \\
+A & 10,660 & $0.981 \pm 0.001$ & $93.6 \pm 0.3$ & $0.60 \pm 0.02$ & $87.4 \pm 1.7$ & $85.4 \pm 2.6$ & False \\
+A & 5,330 & $0.975 \pm 0.003$ & $92.9 \pm 0.6$ & $0.57 \pm 0.03$ & $86.6\pm 2.0$ & $83.4 \pm 3.4$ & True \\
+A & 5,330 & $0.973 \pm 0.002$ & $92.3 \pm 0.4$ & $0.55 \pm 0.02$ & $86.2 \pm 2.4$ & $80.8 \pm 3.8$ & False \\
+B & 1,103 & $0.910 \pm 0.012$ & $85.9 \pm 0.9$ & $0.31 \pm 0.03$ & $72.4 \pm 0.4$ & $66.1 \pm 6.0$ & True \\
+B & 1,103 & $0.901 \pm 0.016$ & $84.6 \pm 1.7$ & $0.28 \pm 0.05$ & $68.2 \pm 3.4$ & $66.3 \pm 5.5$ & False \\
+C & $\sim$10,660 & - & - & 0.58 & 85 & 88 & True \\
+C & $\sim$10,660 & - & - & 0.51 & 82 & 85 & False \\
+C & 1,045 & - & - & 0.33 & 70 & 75 & True \\
+C & 1,045 & - & - & 0.29 & 67 & 71 & False \\
+D & $\sim$8,000 & - & - & 0.55 & - & - & True \\
+D & $\sim$2,00$\underset{~}{0}$ & - & - & 0.45 & - & - & True \\
+\hline\hline
+E & 10,660 & $0.974 \pm 0.001$ & $90.4 \pm 0.3$ & - & $90.6 \pm 0.7$ & $86.5 \pm 0.7$ & True\\
+E & 10,660 & $0.959 \pm 0.006$ & $88.5 \pm 1.1$ & - & $87.6 \pm 1.1$ & $85.9 \pm 4.1$ & False\\
+F & 1,103 & $0.868 \pm 0.015$ & $78.1 \pm 0.9$ & - & $70.8 \pm 3.4$ & $70.6 \pm 4.1$ & True\\
+F & 1,103 & $0.865 \pm 0.011$ & $78.0 \pm 1.2$ & - & $66.9 \pm 3.2$ & $74.5 \pm 4.2$ & False\\\hline\hline
+A & 10,660 & $0.977 \pm 0.002$ & $93.1 \pm 0.4$ & $0.58 \pm 0.01$ & $88.0 \pm 1.1$ & $82.2 \pm 2.8$ & True\\
+A & 10,660 & $0.970 \pm 0.001$ & $92.0 \pm 0.3$ & $0.53 \pm 0.01$ & $86.0 \pm 0.9$ & $79.5 \pm 2.2$ & False\\
+B & 1,103 & $0.902 \pm 0.014$ & $85.2 \pm 1.2$ & $0.29 \pm 0.04$ & $71.5 \pm 1.6$ & $62.8 \pm 5.6$ & True\\
+B & 1,103 & $0.860 \pm 0.017$ & $81.6 \pm 1.2$ & $0.21 \pm 0.02$ & $62.6 \pm 3.0$ & $57.6 \pm 2.7$ & False\\\hline\hline
+E & 10,660 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & True\\
+E & 10,660 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & False\\
+F & 1,103 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & True\\
+F & 1,103 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & False\\\hline\hline
 \end{tabular}
-\caption{\label{tab:05_SN1a} Summary of results for  type-Ia vs.
+\caption{\label{tab:05_SN1a} (Top section) Summary of results for  type-Ia vs.
+non-type-Ia classification with a training fraction of 0.5, 0.25 and 0.052 with comparisons to similar methods in~\cite{Karpenka:2012pm} and~\cite{Newling:2010bp}.
+(Second section) Summary of results for types-I,~-II~and~-III classification.
+(Third section) Summary of results for SPCC early-epoch challenge.
+(Bottom section) Summary of the results for the SPCC early-epoch challenge when classifying between types-I~,~-II~and~-III supernovae.
+The models used are A) Unidirectional LSTM, [16, 16] with 0.5 dropout, B) Unidirectional LSTM, [4] with 0.5 dropout, C) \cite{Karpenka:2012pm}, D) \cite{Newling:2010bp}, E) Bidirectional LSTM, [16, 16] with 0.5 dropout, F) Bidirectional LSTM, [4] with 0.5 dropout.
+Errors on results are the mean and standard deviation values from 5 randomised runs.}
+\end{table*}
+
+%\begin{table}[t!]
+%\begin{subtable}
+%\centering
+%\begin{tabular}{|c | c | c | c | c |} \hline 
+%AUC & Accuracy & F1 & Host z \\ \hline 
+%$ 0.986 \pm  0.001$ &  $  94.8 \pm 0.2\% $ & $  0.64 \pm 0.01  $ & T  \\
+%$ 0.981 \pm  0.002$ &  $  93.6 \pm 0.3\% $ & $  0.60 \pm 0.02  $ & F  \\ \hline
+%\end{tabular}
+%\caption{\label{tab:05_SN1a} Summary of results for  type-Ia vs.
 non-type-Ia classification with a training fraction of 0.5.
 The network is a 2-layer LSTM, each with 16 hidden units and dropout of 0.5.}
-\end{subtable}%
-\begin{subtable}
-\centering
-\begin{tabular}{|c | c | c | c | c |} \hline 
-AUC & Accuracy & Host z  \\ \hline 
-$ 0.971  \pm 0.003 $ &  $  90.4  \pm 0.4 \% $ & T  \\
-$ 0.958 \pm  0.007 $ &  $   88.8 \pm 1.0 \% $ & F \\ \hline
-\end{tabular}
-\caption{\label{tab:05_123} Summary of results for three class categorisation (supernovae types-1, -2 and -3). The model is the same as in Table~\ref{tab:05_SN1a}}
-\end{subtable}
-\end{table}
+%\end{subtable}%
+%\begin{subtable}
+%\centering
+%\begin{tabular}{|c | c | c | c | c |} \hline 
+%AUC & Accuracy & Host z  \\ \hline 
+%$ 0.971  \pm 0.003 $ &  $  90.4  \pm 0.4 \% $ & T  \\
+%$ 0.958 \pm  0.007 $ &  $   88.8 \pm 1.0 \% $ & F \\ \hline
+%\end{tabular}
+%\caption{\label{tab:05_123} Summary of results for three class categorisation (supernovae types-1, -2 and -3). The model is the same as in Table~\ref{tab:05_SN1a}}
+%\end{subtable}
+%\end{table}
 
-One advantage of our approach is that light curve data can be directly inputted to a {\em pre-trained} model to give very fast evaluation ($<1$s) of supernovae type.
+One advantage of our approach is that light curve data can be directly input to a {\em pre-trained} model to give very fast evaluation ($<1$s) of supernovae type.
 In the lower panel of Fig.~\ref{fig:lightcurve} we input the light curve, as a function of time, of a type-Ia supernovae (excluded from training) to the pre-trained 2-layer LSTM model discussed above.
@@
 We also test the same model using a training fraction of 0.25 (around 5000 supernovae), closer to the lower end of the number likely to be followed up for the LSST.
-After 5 randomised runs and training for 200 epochs we obtain an accuracy of $93.3 \pm 0.4\%$, AUC of $0.978 \pm 0.002$ and $F_1 = 0.57 \pm 0.02$.
-The corresponding type-Ia purity and completeness are 85\% and 87\% respectively.
+After 5 randomised runs and training for 200 epochs we obtain an accuracy of $92.9 \pm 0.6\%$, AUC of $0.975 \pm 0.003$ and $F_1 = 0.57 \pm 0.03$.
+The corresponding type-Ia purity and completeness are $86.6\pm2.0\%$ and $83.4\pm3.4\%$ respectively.
 The $F_1$ metric has degraded by $\sim10\%$ for a reduction in data of $50\%$.
@@
 The loss in the latter case would be lower however, and represents increased confidence of the network in its predictions.
-We therefore reject models with severe over-fitting and an increasing cross-entropy loss, at the expense of metrics such as $F_1$, and decrease the model complexity.
+We therefore reject models with severe over-fitting and an increasing cross-entropy loss at the expense of metrics such as $F_1$, and decrease the model complexity.
 
 For a training fraction of  $5.2\%$ we find a single-layer LSTM network, with 4 hidden units, and dropout of 0.5 satisfies this criteria.
-For 5 randomised runs, training for 200 epochs, we obtain a classification accuracy of $85.9 \pm 1.2$\%, AUC of $0.910 \pm 0.012$ and $F_1 = 0.31 \pm 0.03$.
-The corresponding type-Ia purity and completeness are 72\% and 66\% respectively.
+For 5 randomised runs, training for 200 epochs, we obtain a classification accuracy of $85.9 \pm 0.9$\%, AUC of $0.910 \pm 0.012$ and $F_1 = 0.31 \pm 0.03$.
+The corresponding type-Ia purity and completeness are $72.4\pm0.4\%$ and $66.1\pm6.0\%$ respectively.
+
+It is difficult to directly compare the results from the SPCC directly with this work, but using post-SPCC data, we obtain results which are competitive with previous approaches.
+The analyses by~\cite{Karpenka:2012pm} and~\cite{Newling:2010bp} are easier to compare.
+Along with~\cite{Lochner:2016hbn} these employ a two-step process, where features are first extracted by various methods before machine learning classification.
+The results obtained for similar sized training sets are comparable as can be seen in the top section of Table~\ref{tab:05_SN1a}.
+When using half the dataset to train on we get a higher $F_1$ value, $F_1=0.64$ compared to $F_1=0.58$ in~\cite{Karpenka:2012pm}.
+The value in~\cite{Newling:2010bp} is also similar given that the sample size is smaller.
+For a smaller sample training set of 5.2\% of all the data we again perform similarly to~\cite{Karpenka:2012pm} but under perform compare to~\cite{Newling:2010bp} taking into account the slightly larger sample size in the latter case.
+
+In the middle section of Table~\ref{tab:05_SN1a} the three class categorisation is shown.
+There is no available data for comparison of this problem, but compared to classification between type-Ia vs.
+non-type-Ia, bidirectional recurrent neural networks do well.
+The AUC and accuracy remain high, still above 90\% when the host-$z$ is included using a training fraction of 0.5
+
+Finally the bottom section of Table~\ref{tab:05_SN1a} shows the results of the early-epoch challenge from SPCC.
+Here only the data before the night of the sixth observation with $S/N>4$ for each light-curve can be used - a great reduction from the use of the full light-curve.
+We do surprisingly well in this case obtaining an accuracy of $93.1\pm0.4\%$, AUC of $0.977\pm0.002$ and an $F_1=0.58\pm 0.01$ with a training fraction of 0.5 and including host-$z$.
+These values are not far from those obtained using the whole light-curve and are equivalent to the full results of~\cite{Karpenka:2012pm}.
+The results are not as good with a training fraction of 0.052, but still comparable to our results using the whole light-curve.
+The network trained on the partial light-curves does better than suggested from feeding the early-epoch light-curve through a network trained on the full sequence.
+This is due to the later parts of the light-curve influencing the weights of the network whilst training.
+Training on only the initial part of the light-curve optimises the network weights such that early sequence features have more effect, resulting in better accuracy, AUC and $F_1$ values than expected.
 
@@
 Machine learning methodology has previously been applied to SPCC classification~\cite{Newling:2010bp, Karpenka:2012pm, Lochner:2016hbn}.
-These employ a two-step process, where features are first extracted by some method before classification.
-In our approach light curves are used directly as inputs to a recurrent neural network, which is able to learn information from the sequence of observations.
+Instead of performing feature extraction before classification, our approach uses the light-curves directly as inputs to a recurrent neural network, which is able to learn information from the sequence of observations.
 
-Using post-SPCC data, we obtain results which are competitive with previous approaches.
-Although we have trained the network on the cross-entropy loss and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-z), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.65).
+Although we have trained the network on the cross-entropy loss and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-$z$), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.64).
 Recurrent neural networks  therefore compare well with other methods when a larger training  set is available.
@@
 This is possible since the light curve can be fed to the model directly without needing any feature extraction.
+Moreover, we have found that training a network only on the early epoch light-curve data results in a better early-time predictor than using a network trained on entire light-curve data.
 
@@
 Finally, to improve performance in the small data regime one can use {\em transfer learning.} Here, a more complex network is pre-trained on simulations or existing data from other surveys, then the weights of the network are fine-tuned on the new, smaller dataset.
-The simulated SPCC data used in this work are based on the DES instrument, so it would be interesting to apply transfer learning to real DES data.
+The simulated SPCC data used in this work are based on the DES instrument, and we are applying transfer learning to real DES data for publication in future work.
 
@@
 
+\bibitem[{Medsker \& Jain(1999)}]{medsker1999recurrent}
+Medsker, L., \& Jain, L.~C. 1999, Recurrent neural networks: design and
+  applications (CRC press)
+
 \bibitem[{Newling {et~al.}(2011)}]{Newling:2010bp}

### 2017-01-25 arch.py

@@ -35,8 +35,8 @@ def analyse(filename):
 		print('classifier: ', m)
 
 		with open("arch/"+filename+"_"+m+".tex", "w") as f:
-			f.write("\\begin{tabular}{|" + " | ".join(["c"] * 9) + "|} \hline \n")
-			f.write(" & ".join(['AUC', 'Accuracy', 'F1', 'z', 'Class', 'h', 'B']) + " \\\\ \hline \n")
+			f.write("\\begin{tabular}{|" + " | ".join(["c"] * 10) + "|} \hline \n")
+			f.write(" & ".join(['AUC', 'Accuracy', 'F1', 'z', 'Class', 'h', 'B', 'Dropout', 'Detection']) + " \\\\ \hline \n")
 			for row in sorted_results:
 				p = row[2]
 				acc_mean = np.mean(p['acc'])
@@ -51,6 +51,10 @@ def analyse(filename):
 				prec_std = [ "$ {:.3f}".format(np.std(prec[:,i])) for i in range(0,prec.shape[1])]
 				rec_mean = [ "$ {:.3f}".format(np.mean(rec[:,i])) for i in range(0,prec.shape[1])]
 				rec_std = [ "$ {:.3f}".format(np.std(rec[:,i])) for i in range(0,prec.shape[1])]
+				#prec_mean = [ "$ {:.3f}".format(np.mean(prec[:,i])) for i in range(0,prec.shape[1])]
+				#prec_std = [ "$ {:.3f}".format(np.std(prec[:,i])) for i in range(0,prec.shape[1])]
+				#rec_mean = [ "$ {:.3f}".format(np.mean(rec[:,i])) for i in range(0,prec.shape[1])]
+				#rec_std = [ "$ {:.3f}".format(np.std(rec[:,i])) for i in range(0,prec.shape[1])]
 				auc = "$ {:.3f}".format(auc_mean) + ' \pm ' + "{:.3f} $".format(auc_std)
 				acc = "$ {:.3f}".format(acc_mean) + ' \pm ' + "{:.3f} $".format(acc_std)
 				f1 = "$ {:.2f}".format(f1_mean) + ' \pm ' + "{:.2f} $".format(f1_std)
@@ -62,8 +66,8 @@ def analyse(filename):
 					host = 'F'
 				else:
 					host = 'T'
-				f.write(" & ".join([row[3], auc, acc, f1, host, p['rnn'], str(p['layers']), bi, str(p['dropout'])]) + " \\\\\n")
[…]

### 2017-01-25 dataset.py

@@ -79,7 +79,7 @@ def to_categorical(y, nb_classes=None):
 		Y[i, y[i]] = 1.
 	return Y
 
-def load_data(path="data/unblind_nohostz", classifier=sn1a_classifier, test_fraction=0.2, nb_augment=1, seed=None):
+def load_data(path="data/unblind_nohostz", classifier=sn1a_classifier, test_fraction=0.2, nb_augment=1, seed=None, detection="All"):
 	'''
 	Loads data from the files produced by preprocess.py. Returns the data as numpy arrays which can be used in the
         model.
@@ -90,6 +90,7 @@ def load_data(path="data/unblind_nohostz", classifier=sn1a_classifier, test_frac
           1-test_fraction is the fraction of the data used to train the network
         * nb_augment is the number of random augmentations of data to use
 	* seed is an integer required to prevent different random values
+	* detection is a string to choose whether to use only light curve values upto detection tag. Should be set to "Test" or "Both" to take only the inital values ("All" is default)
 	* nb_samples is the number of different supernovae
 	* sequence_length is the length of the sequences in each supernova event (all equal using pad_sequence())
 	* output_dim is the number of elements in the results (12 with no host and 13 with host)
@@ -140,6 +141,9 @@ def load_data(path="data/unblind_nohostz", classifier=sn1a_classifier, test_frac
 	data = []
 	training_idx = []
 	test_idx = []
+	first_ids = []
+	first_idx = []
+	detection_idx = []
 	idx = 0
 
 	ids_train_permute = []
@@ -157,11 +161,21 @@ def load_data(path="data/unblind_nohostz", classifier=sn1a_classifier, test_frac
 			for row in reader:
 				id = int(row[0])
 				if id != last_id:
+					if id not in first_ids:
+						first_ids.append(id)
+						if detection == 'Both':
+							first_idx.append(idx)	
+							detection_idx.append(int(row[1]))
+						if detection == 'Test':
+							if id in ids_test:
+								first_idx.append(idx)	
+								detection_idx.append(int(row[1]))
 					if not first_time:
 						data.append(data_sequence)
 						if id in ids_train:
 							training_idx.append(idx)
 							ids_train_permute.append(last_id)
+								
 						else:
 							test_idx.append(idx)
 							ids_test_permute.append(last_id)
[…]

### 2017-01-25 preprocess.py

@@ -141,6 +141,7 @@ def parser_last(filename):
 	* filename is a string containing the path to the supernovae light curve data
 	* survey is a string containing the survey name
 	* snid is an integer containing the supernova ID
+	* detection_index is an integer containing the last sequence index for challenge 2
 	* ra is a float containing the RA of the supernova
 	* dec is a float containing the Dec of the supernova
 	* mwebv is a float describing the dust extinction
@@ -152,7 +153,7 @@ def parser_last(filename):
 	* obs is a sequence of arrays each element containing [time since first observation,fluxes in each colourband,flux errors in each colourband]
 	- Used in __main__() to read in the data
 	'''
-	survey = snid = ra = dec = mwebv = hostid = hostz = spec = sim_type = sim_z = None
+	survey = snid = detection = ra = dec = mwebv = hostid = hostz = spec = sim_type = sim_z = None
 	obs = []
 	g = r = i = z = 0
 	g_error = r_error = i_error = z_error = 0
@@ -183,6 +184,8 @@ def parser_last(filename):
 					sim_type = s[1].split('SN Type =')[1].split(',')[0].strip()
 				elif s[0] == 'SIM_REDSHIFT':
 					sim_z = float(s[1])
+				elif s[0] == 'DETECTION':
+					detection_index = len(obs)
 				elif s[0] == 'OBS':
 					o = s[1].split() 
 					if first_obs is None:
@@ -200,7 +203,7 @@ def parser_last(filename):
 						z = float(o[3])/flux_norm
 						z_error = float(o[4])/flux_norm
 					obs.append([(float(o[0]) - first_obs)/time_norm] + [g,r,i,z] + [g_error,r_error,i_error,z_error])
-	return survey, snid, sn_type, sim_type, sim_z, ra, decl, mwebv, hostid, hostz, spec, obs
+	return survey, snid, detection_index, sn_type, sim_type, sim_z, ra, decl, mwebv, hostid, hostz, spec, obs
 
 def parser_spline(filename):
 	'''
@@ -209,6 +212,7 @@ def parser_spline(filename):
 	* filename is a string containing the path to the supernovae light curve data
 	* survey is a string containing the survey name
 	* snid is an integer containing the supernova ID
[…]

### 2017-01-25 test.py

@@ -1,0 +1,91 @@
+#from dataset import load_data
+#sn1a_classifier = {'1':0, '2':1, '3':1, '21':1, '22': 1, '23':1, '32': 1, '33':1}
+#(X_train, X_train_reverse, Y_train, ids_train), (X_test, X_test_reverse, Y_test, ids_test), (length_train, length_test, sequence_len, output_dim, nb_classes) = load_data(path="data/unblind_nohostz", test_fraction=0.5, classifier=sn1a_classifier, nb_augment=1, detection="Both")
+
+import glob
+import json
+import numpy as np
+def analyse(filename):
+
+	results = []
+	maps = []
+
+	#for f in glob.glob('arch/*.txt'):
+	#SN1a and 123
+	for f in ['arch/lstm_0.5_USUCTL.txt', 'arch/lstm_0.5_RBAFYR.txt', 'arch/lstm_0.75_IEXIOJ.txt', 'arch/lstm_0.75_YZIMTD.txt', 'arch/lstm_0.948_HTDJUB.txt', 'arch/lstm_0.948_HHGQPA.txt', 'arch/lstm_0.5_ZQVVVM.txt', 'arch/lstm_0.5_QDJMVX.txt', 'arch/lstm_0.948_WQQXYR.txt', 'arch/lstm_0.948_MOVTRG.txt']:
+	#SN1a and 123 early epoch
+	#for f in ['arch/info_0.5_OMEQFR.txt', 'arch/info_0.5_NVMIOZ.txt', 'arch/info_0.5_WSXTAT.txt', 'arch/info_0.948_KUDLXC.txt', 'arch/info_0.948_KDXFCS.txt', 'arch/info_0.948_PFNMKJ.txt', 'arch/info_0.5_CLKQZA.txt', 'arch/info_0.948_RSOVMJ.txt']:
+		print f
+		with open(f) as data_file:   
+			p = json.load(data_file)
+			if 'auc' in p:
+				try:
+					results.append([np.mean(p['f1']), p['classifier'][1], p, f])
+					maps.append(p['classifier'][1])
+				except:
+					results.append([np.mean(p['f1']), p['mapping'][1], p, f])
+					maps.append(p['mapping'][1])
+	for m in list(set(maps)):
+
+		map_results = [r for r in results if r[1] == m]
+		sorted_results = sorted(map_results, key=lambda x: -x[0])
+
+		print('classifier: ', m)
+
+		with open("arch/output_" + m + ".tex", "w") as f:
[…]

### 2017-01-25 train.py

@@ -99,7 +99,7 @@ def build_model(nb_hidden, nb_classes, dropout, sequence_len, output_dim, bidire
 
 def train(batch_size=10, dropout=0.5, nb_hidden=[16, 16], path="data/unblind_nohostz", test_fraction=0.5, classifier=sn1a_classifier, nb_epoch=200, nb_augment=5, 
 	bidirectional=False, rnn_type='LSTM',feedback=1, optimizer='adam', activation='tanh', save_model=False, plot_loss=False, plot_data=[], 
-        filename='current', consensus=True):
+        filename='current', consensus=True, detection='All'):
 	'''
 	Main training module. The data is read and the neural network built and the data fed into the neural network.
 	The analysis metrics are also calculated here.
@@ -126,6 +126,8 @@ def train(batch_size=10, dropout=0.5, nb_hidden=[16, 16], path="data/unblind_noh
 	  is appended if more than one random augmentation is used
 	* filename is a string containing the save name for plots and model properties 
 	* consensus is a boolean which when true mean pools the predictions from each step
+	* detection is a string to choose whether to use only light curve values upto detection tag. Should be set to "Test"
+          or "Both" to take only the inital values ("All" is default)
 	* acc is a float containing the final accuracy
 	* auc is a float containing the final area under the Receiver Operating Characteristic curve
 	* prec is the final precision
@@ -183,7 +185,7 @@ def train(batch_size=10, dropout=0.5, nb_hidden=[16, 16], path="data/unblind_noh
 				self.plot_data[len(self.plot_data)-1] = [self.loss, self.val_loss, self.acc, self.val_acc]
 				loss_plotter(self.plot_data, self.filename)
 
-	(X_train, X_train_reverse, Y_train, ids_train), (X_test, X_test_reverse, Y_test, ids_test), (length_train, length_test, sequence_len, output_dim, nb_classes) = load_data(path=path, test_fraction=test_fraction, classifier=classifier, nb_augment=nb_augment)
+	(X_train, X_train_reverse, Y_train, ids_train), (X_test, X_test_reverse, Y_test, ids_test), (length_train, length_test, sequence_len, output_dim, nb_classes) = load_data(path=path, test_fraction=test_fraction, classifier=classifier, nb_augment=nb_augment, detection=detection)
 
 	model_root = 'save/' + filename
 	

### 2017-01-25 utils.py

@@ -44,11 +44,13 @@ def loss_plotter(history, filename, plot_legend=True):
 			ax1.semilogy(history[i][1],color='blue')
 			ax2.semilogy(1 - np.array(history[i][2]),color='green')
 			ax2.semilogy(1 - np.array(history[i][3]),color='blue')
+
 		else:
 			a, = ax1.semilogy(history[i][0],color='green')
 			b, = ax1.semilogy(history[i][1],color='blue')
 			c, = ax2.semilogy(1 - np.array(history[i][2]),color='green')
 			d, = ax2.semilogy(1 - np.array(history[i][3]),color='blue')
+
 
 	if plot_legend:
 		ax1.legend([a,b],['Train','Test'],loc='best')

### 2017-01-31 paper/sn_submission_v3/Referee reply.txt
Paper update

@@
 
-There is possibly a way to improve the model further, but this is not certain.
-One of our quarms about trying to improve the model further is that the way RNNs work mean that including the information from after the early epoch challenge skews the overall weights of the network meaning that the early epoch weights are less well known compared to training using only the early epoch challenge data.
+There is possibly ways to improve the model further, but this is not certain.
+One reason that improving the model further may not actually help is that, due to the way RNNs work, including the information from after the early epoch challenge skews the overall weights of the network meaning that the early epoch weights are less well known compared to training using only the early epoch challenge data.
+Since this is mentioned in the text we have not expanded further.
 
@@
 
+We have added (new) Fig.2 which contains three subplots.
+The lower left subplot is simply the total number of days of observation of the light-curve vs.
+the number of grouped-time data vector elements these are split into.
+This shows that when there are more days of observation the light-curve data is grouped into longer data vectors.
+The upper subplot shows the distribution of the total number of days of observation of the light-curve with min, max, mean and median values indicated and the lower right subplot shows the distribution of the number of elements in the grouped-time data vector.
+We have added this information to the text.
+
 (3) *ANY* discussion about error-bars will be helpful in letting readers decide if
@@
 
-The errors in the results are due to running the training 5 times with different random selections of training data (and random augmentations of data).
-This is mentioned in the results section already.
+The errors in the results are due to training 5 times with different random selections of training data (and random augmentations of data).
+This is mentioned in the results section already, but we have expanded on it in the same paragraph.
 
@@
 
-To do...
+We ran the training removing 10% of the known points from each filter before grouping the times and randomly augmenting.
+The equivalent of the top line in the table with 10% of the points removed gives results very close to the results in the second line of the table, i.e.
+there is a slight degredation but this is to be expected, and the results are possibly better than we expected when performing this proceduce.
+We have added this to the results section.
 
@@
 This is already explicitly described in the text.
+We feel this short description helps with the understanding of the process of training the network, but does not need expanding.
 
@@
 
-They are already in the table in every case - this a train fraction of 0.052, not 0.052%.
+They are already in the table in every case - 1,045 is equivalent to a train fraction of 0.052 (not 0.052%).
 
@@
 
-The early-epoch challenge is for SN1a only and not 123. We have given that a go and added the results to the table.
+The early-epoch challenge is for SN1a only and not 123.
+We have given 123 a go and added the results to the table.
+The difference between 123 categorisation using the full light-curve and the early-epoch data is very similar to the difference between the SN1a full light-curve and early-epoch data.
+The results are not as good, but still surprisingly high considering the length of the vectors.
 
@@
 As it can be seen the errors on both of the bottom lines of the second section are ~4%, approximately the same as the difference between the means.
+The same is true for the 123 categorisation in the early-epoch challenge where the completeness without host-z is higher than with host-z, but the error on the completeness is also very large.
 
@@
 
-To do..
+We have highlighted the result in the abstract as well as separated the paragraph in the conclusion section.
+The main point of the paper is to describe the network architecture and it's use in classification, rather than achieve the best results possible.
+It is somewhat fortunate that we did as well as we did on the early-epoch challenge, but we don't want to make that the main focus of the paper.
 
@@
 It is possible to data mine images and pipe this into an RNN.
-We could do this for a single filter, we could do this for many filters.
+We could do this for a single filter or we could do this for many filters.
 If one filter was missing then the input could be augmented as done here and the other raw images convolved to extract features.
-This is identical to what is done here, but with a huge amount more information.
+The principal is identical to what is done here, but with a huge amount more information.
+We have not changed the text here as it is an interesting prospect which others (as well as ourselves) might like to consider.
 
@@
 
-What is different between differences in inputs and noise? They are the same as far as we can tell
+We have changed this to describe that over-fitting occurs when a network learns about relations between the inputs and outputs in the training data which aren't present in the test data.
 
@@
 
-Done.
+We have changed the scale to show between 3e-2 and 4e-1 therefore reducing the unused white-space.
 
@@
 
-We are not sure how it's supposed to feel?
+We are not sure what the problem is here.
 
@@
 
-We have changed this
+We have changed this.

### 2017-01-31 paper/sn_submission_v3/ms.tex
Paper update

@@
 For a training size of 50\% of the representational SPCC dataset (around $10^4$ supernovae) we obtain a type-Ia vs.
-non-type-Ia classification accuracy of 94.8\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$.
+non-type-Ia classification accuracy of 94.7\%, an area under the Receiver Operating Characteristic curve AUC of 0.986 and a SPCC figure-of-merit $F_1=0.64$.
+When using only the data for the early-epoch challenge defined by the SPCC we achieve a classification accuracy of 93.1\%, AUC of 0.977 and $F_1=0.58$, results almost as good as with the whole light-curve.
+By employing bidirectional neural networks we can acquire impressive classification results between supernovae types -I,~-II and~-III at an accuracy of 90.4\% and AUC of 0.974.
 We also apply a pre-trained model to obtain classification probabilities as a function of time, and show it can give early indications of supernovae type.
@@
 \centering
-\includegraphics[width=88mm, angle=0]{f1.pdf}
+\includegraphics[width=83mm, angle=0]{f1.pdf}
 \caption{\label{fig:network} Bidirectional recurrent neural network for sequence classification.
@@
 
+In Fig.~\ref{fig:violinplot} we show how the length of the grouped-time data vector is related to the duration of the light-curve.
+The bottom left subplot shows that more total number of day since the beginning of observation of the light-curve results in a greater number of grouped time elements in the vector.
+The upper subplot shows the distribution of observation lengths in the SPCC data varies significantly with two distinct peaks.
+These are grouped into an average of 40-3element data vectors as can be seen in the bottom right subplot.
+
+\begin{figure}
+\centering
+\includegraphics[width=84mm, angle=0]{f2.pdf}
+\caption{\label{fig:violinplot}.
+(Top) Distribution of the total number of days for each light-curve with the minimum, maximum, mean and median values indicated.
+(Bottom right) Distribution of the number of elements in the grouped time vector with the minimum, maximum, mean and median values indicated.
+(Bottom left) The trend showing that more days in the light-curve result in longer group time vectors.
+ }
+\end{figure}
+
 Observations are of the form in Table~\ref{tab:augment}, where any missing values are denoted by a dash.
@@
 \centering
-\includegraphics[width=88mm, angle=0]{f2.pdf}
+\includegraphics[width=84mm, angle=0]{f3.pdf}
 \caption{\label{fig:lightcurve} (Top) Example light curve in the 4 $g,r,i,z$ bands for SN ID 551675 (a type-Ia) in the Supernovae Photometric Classification Challenge data~\cite{Kessler:2010wk}.
@@
 We denote these as  `SN1a' and  `123' respectively.
-We also attempt to categorise between type-Ia and non-type-Ia supernovae using only the first six observations with $S/N>4$ and the data taken on the night of the sixth observation as described in~\cite{Kessler:2010wk}.
+We also attempt the first two problems using only the first six observations with $S/N>4$ and the data taken on the night of the sixth observation as described in~\cite{Kessler:2010wk}.
 
@@
 \centering
-\includegraphics[width=88mm, angle=0]{f3.pdf}
+\includegraphics[width=84mm, angle=0]{f4.pdf}
 \caption{\label{fig:loss} (Left) Training loss (green) vs.
@@
 We therefore need to be careful about {\em over-fitting}.
-Over-fitting arises when the network fits noise between the inputs and outputs of the training data, which do not exist in the test data.
+Over-fitting arises when the network learns about relations between the inputs and outputs of the training data, which do not exist in the test data.
 It can typically be detected by comparing the loss of the training and test data.
@@
 The inclusion of host galaxy photo-$z$ marginally improves the classifier performance.
+The 1$\sigma$ errors quoted in the table are the result of 5 runs where the training data is randomly chosen (and so different) each time.
+Some random choice of the set of light-curves are more effective for training the network than others, but it is extremely difficult to optimise this.
+
+To test the robustness of the time-grouping method we remove 10\% of the known filter values (and/or their errors) before grouping the data into a single vector and randomly augmenting the missing values.
+After training we find there is a small degradation in the results, i.e.
+for a training fraction of 0.5 using a deep 2-layer, unidirectional network with 16 hidden units, a dropout of 0.5 and including the photo-$z$ information the obtained results are very similar to the second line in Table~\ref{tab:05_SN1a}.
+This shows that a reduction in 10\% of the points is similar to the omission of the photo-$z$ data and therefore the data augmentation method is extremely robust.
 
@@
 B & 1,103 & $0.860 \pm 0.017$ & $81.6 \pm 1.2$ & $0.21 \pm 0.02$ & $62.6 \pm 3.0$ & $57.6 \pm 2.7$ & False\\\hline\hline
-E & 10,660 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & True\\
-E & 10,660 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & False\\
-F & 1,103 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & True\\
-F & 1,103 & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & $*** \pm ***$ & False\\\hline\hline
+A & 10,660 & $0.960 \pm 0.006$ & $87.9 \pm 0.9$ & - & $86.4 \pm 0.8$ & $84.4 \pm 3.5$ & True\\
+A & 10,660 & $0.948 \pm 0.002$ & $86.8 \pm 0.3$ & - & $84.1 \pm 1.1$ & $83.7 \pm 1.4$ & False\\
+B & 1,103 & $0.851 \pm 0.013$ & $76.8 \pm 1.3$ & - & $64.7 \pm 3.8$ & $71.0 \pm 4.1$ & True\\
+B & 1,103 & $0.819 \pm 0.010$ & $74.2 \pm 1.0$ & - & $58.1 \pm 3.8$ & $73.6 \pm 6.6$ & False\\\hline\hline
 \end{tabular}
@@
 The reason for this apparent discrepancy is that the accuracy, say, simply takes the maximum value of the softmax output layer.
-For example, a 2 class problem with output probabilities [0.6, 0.4] and target [1, 0] has the same accuracy as one with output probabilities [0.8, 0.2].
+For example, a 2-class problem with output probabilities [0.6, 0.4] and target [1, 0] has the same accuracy as one with output probabilities [0.8, 0.2].
 The loss in the latter case would be lower however, and represents increased confidence of the network in its predictions.
@@
 
-In the middle section of Table~\ref{tab:05_SN1a} the three class categorisation is shown.
+In the second section of Table~\ref{tab:05_SN1a} the three class categorisation is shown.
 There is no available data for comparison of this problem, but compared to classification between type-Ia vs.
 non-type-Ia, bidirectional recurrent neural networks do well.
-The AUC and accuracy remain high, still above 90\% when the host-$z$ is included using a training fraction of 0.5
+The AUC and accuracy remain high, still above 90\% when the host-$z$ is included using a training fraction of 0.5.
+Using a smaller training fraction of 0.052, the results are worsened similar to the two class categorisation in the top section of Table~\ref{tab:05_SN1a}.
 
-Finally the bottom section of Table~\ref{tab:05_SN1a} shows the results of the early-epoch challenge from SPCC.
+The third section of Table~\ref{tab:05_SN1a} shows the results of the early-epoch challenge from SPCC.
 Here only the data before the night of the sixth observation with $S/N>4$ for each light-curve can be used - a great reduction from the use of the full light-curve.
@@
 Training on only the initial part of the light-curve optimises the network weights such that early sequence features have more effect, resulting in better accuracy, AUC and $F_1$ values than expected.
+
+Finally, the bottom section of Table~\ref{tab:05_SN1a} has the results of the three class categorisation when using the early-epoch data.
+The results are similar to the difference between the full light-curve and early-epoch data SN1a categorisation when comparing with the full light-curve 123 categorisation.
+It should be noted that the bidirectional network used for the 123 categorisation using the full light-curve revealed sizeable over-fitting when using the early-epoch data and so a unidirectional network was used instead.
 
@@
 This is possible since the light curve can be fed to the model directly without needing any feature extraction.
-Moreover, we have found that training a network only on the early epoch light-curve data results in a better early-time predictor than using a network trained on entire light-curve data.
+
+Most interestingly, we have found that training a network only on the early epoch light-curve data results in a better early-time predictor than using a network trained on entire light-curve data.
+Our results using only the early-epoch data are close to those using the entire light-curve data for both SN1a and 123 categorisation with both large and small training fractions.
 

### 2017-02-06 paper/sn_submission_v3/Referee reply.txt
v3 Paper submission

@@
+We wish to thank the referee for their points and we have done our best to expand on each of them fully.
+
 (1) The work as presented is in almost acceptable form. The part about using just the
@@
 
-There is possibly ways to improve the model further, but this is not certain.
+There could be possibly ways to improve the model further, but this is not a certainty.
 One reason that improving the model further may not actually help is that, due to the way RNNs work, including the information from after the early epoch challenge skews the overall weights of the network meaning that the early epoch weights are less well known compared to training using only the early epoch challenge data.
@@
 This means the padding has no effect on the results, it just maintains the correct length vectors.
-This is already explicitly described in the text.
+This is already described in the text.
 We feel this short description helps with the understanding of the process of training the network, but does not need expanding.
@@
 The difference between 123 categorisation using the full light-curve and the early-epoch data is very similar to the difference between the SN1a full light-curve and early-epoch data.
-The results are not as good, but still surprisingly high considering the length of the vectors.
+The results are not as good, but still surprisingly high considering the length of the data.
 
@@
 
-To do...
+We have added a paragraph stating the numbers of the highest average F_1 method (Sako) from the SPCC results paper.
+We have also said that they other methods performed similarly but they cannot be compared well to our results.
 
@@
 The principal is identical to what is done here, but with a huge amount more information.
-We have not changed the text here as it is an interesting prospect which others (as well as ourselves) might like to consider.
+We have not changed the text here as it is an interesting prospect which others (as well as ourselves) might like to consider, but would need considerable work to go into detail.
 

### 2017-02-06 paper/sn_submission_v3/ms.tex
v3 Paper submission

@@
 where $i,j$ run over the class labels, $t_{i,j}$ are the targets for each class (either 0 or 1) and $p_{i,j}$ are the predicted probabilities.
-Back-propogation takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.
+Back-propagation takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.
 
@@
 
-In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, consisting of 21319 simulated supernove light curves.
+In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, consisting of 21319 simulated supernova light curves.
 Each supernovae sample consists of a time series of flux measurements, with errors, in the $g,r,i,z$ bands (one band for each timestep), along with the position on the sky and dust extinction.
@@
 The upper subplot shows the distribution of observation lengths in the SPCC data varies significantly with two distinct peaks.
-These are grouped into an average of 40-3element data vectors as can be seen in the bottom right subplot.
+These are grouped into an average of 40-element data vectors as can be seen in the bottom right subplot.
 
@@
 
-It is difficult to directly compare the results from the SPCC directly with this work, but using post-SPCC data, we obtain results which are competitive with previous approaches.
+It is difficult to directly compare the results from the SPCC challenge in~\cite{Kessler:2010qj} with this work since the figure of merit is quoted as a function of redshift and a non-representative set of light-curves was originally used.
+In~\cite{Kessler:2010qj} the method of~\cite{2008AJ....135..348S} had the highest average $F_1\sim0.79$, with 76\% purity and the 96\% completeness.
+This is a, somewhat, confusing average as $F_1\sim 0.4$ at a redshift $z\sim0.1$ up to $F_1\sim1$ at $z\sim0.9$.
+Other methods performed similarly.
+
+It is better to consider comparison with other methods using post-SPCC data, for we obtain results which are competitive with previous approaches.
 The analyses by~\cite{Karpenka:2012pm} and~\cite{Newling:2010bp} are easier to compare.
@@
 
+\bibitem[Sako et al.(2008)]{2008AJ....135..348S} Sako, M., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 348-373
+
 \bibitem[{Srivastava {et~al.}(2014)Srivastava, Hinton, Krizhevsky, Sutskever,

### 2017-02-09 paper/sn_submission_v3/ms.tex
v3 Paper submission

@@
 D & $\sim$2,00$\underset{~}{0}$ & - & - & 0.45 & - & - & True \\
+E & 1,103 & $0.94 \pm 0.03$ & - & - & - & - & True \\
+E & 1,103 & $0.89 \pm 0.53$ & - & - & - & - & False \\
+E & 1,103 & - & - & - & 90 & 85 & True\\
+E & 1,103 & - & - & - & 87 & 90 & True\\
 \hline\hline
-E & 10,660 & $0.974 \pm 0.001$ & $90.4 \pm 0.3$ & - & $90.6 \pm 0.7$ & $86.5 \pm 0.7$ & True\\
-E & 10,660 & $0.959 \pm 0.006$ & $88.5 \pm 1.1$ & - & $87.6 \pm 1.1$ & $85.9 \pm 4.1$ & False\\
-F & 1,103 & $0.868 \pm 0.015$ & $78.1 \pm 0.9$ & - & $70.8 \pm 3.4$ & $70.6 \pm 4.1$ & True\\
-F & 1,103 & $0.865 \pm 0.011$ & $78.0 \pm 1.2$ & - & $66.9 \pm 3.2$ & $74.5 \pm 4.2$ & False\\\hline\hline
+F & 10,660 & $0.974 \pm 0.001$ & $90.4 \pm 0.3$ & - & $90.6 \pm 0.7$ & $86.5 \pm 0.7$ & True\\
+F & 10,660 & $0.959 \pm 0.006$ & $88.5 \pm 1.1$ & - & $87.6 \pm 1.1$ & $85.9 \pm 4.1$ & False\\
+G & 1,103 & $0.868 \pm 0.015$ & $78.1 \pm 0.9$ & - & $70.8 \pm 3.4$ & $70.6 \pm 4.1$ & True\\
+G & 1,103 & $0.865 \pm 0.011$ & $78.0 \pm 1.2$ & - & $66.9 \pm 3.2$ & $74.5 \pm 4.2$ & False\\\hline\hline
 A & 10,660 & $0.977 \pm 0.002$ & $93.1 \pm 0.4$ & $0.58 \pm 0.01$ & $88.0 \pm 1.1$ & $82.2 \pm 2.8$ & True\\
@@
 (Bottom section) Summary of the results for the SPCC early-epoch challenge when classifying between types-I~,~-II~and~-III supernovae.
-The models used are A) Unidirectional LSTM, [16, 16] with 0.5 dropout, B) Unidirectional LSTM, [4] with 0.5 dropout, C) \cite{Karpenka:2012pm}, D) \cite{Newling:2010bp}, E) Bidirectional LSTM, [16, 16] with 0.5 dropout, F) Bidirectional LSTM, [4] with 0.5 dropout.
+The models used are A) Unidirectional LSTM, [16, 16] with 0.5 dropout, B) Unidirectional LSTM, [4] with 0.5 dropout, C) \cite{Karpenka:2012pm}, D) \cite{Newling:2010bp}, E) \cite{Lochner:2016hbn}~SALT2 fits averaged over machine learning architecture F) Bidirectional LSTM, [16, 16] with 0.5 dropout, G) Bidirectional LSTM, [4] with 0.5 dropout.
 Errors on results are the mean and standard deviation values from 5 randomised runs.}
@@
 It is difficult to directly compare the results from the SPCC challenge in~\cite{Kessler:2010qj} with this work since the figure of merit is quoted as a function of redshift and a non-representative set of light-curves was originally used.
-In~\cite{Kessler:2010qj} the method of~\cite{2008AJ....135..348S} had the highest average $F_1\sim0.79$, with 76\% purity and the 96\% completeness.
+In~\cite{Kessler:2010qj} the method of~\cite{2008AJ....135..348S} had the highest average $F_1$, with 79\% purity and 96\% accuracy.
 This is a, somewhat, confusing average as $F_1\sim 0.4$ at a redshift $z\sim0.1$ up to $F_1\sim1$ at $z\sim0.9$.
@@
 For a smaller sample training set of 5.2\% of all the data we again perform similarly to~\cite{Karpenka:2012pm} but under perform compare to~\cite{Newling:2010bp} taking into account the slightly larger sample size in the latter case.
+In~\cite{Lochner:2016hbn} using the SALT2 fits provided the best average AUC over a range of machine learning techniques.
+By imposing a purity of 90\% a completeness of 85\% was achieved while requiring a completeness of 90\% reveals a corresponding purity of 85\%.
 
